### 2018-10-18 notebooks/article_notebook.Rmd
updated last figure to include n

---
title: "Article data analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Dependencies
This notebook is using the following packages:

- dplyr

- ggplot2

- purrr
```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
```

```{r Import data, include = FALSE}

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

two_reads <- read.table("../data/no_of_reads.txt",
                          sep = "\t",
                          header = F,
                          stringsAsFactors = F) %>%
  rename("ref" = V1,
         "reads" = V2) %>%
  mutate(ref = gsub(".+/(.*?)\\.fastq\\.gz", "\\1", ref),
         type = ifelse(grepl("R1", ref) == TRUE, "R1", "R2")) %>%
  mutate(ref = gsub("(.*?)_L00[0-9]_R[1-2]_00[0-9]","\\1", ref)) %>%
  group_by(ref) %>%
  mutate(id = n()) %>%
  {. ->> four_reads} %>%
  filter(id == 2) %>%
  spread(type, reads) %>%
  rename("R1-1" = R1,
         "R2-1" = R2) %>%
  mutate("R1-2" = NA,
         "R2-2" = NA) %>%
  select(-id) %>%
  mutate_all(funs(as.character(.)))

four_reads <- four_reads %>%
  filter(id == 4) %>%
  group_by(ref) %>%
  mutate(id = 1:n()) %>%
  spread(id, reads) %>%
  summarise_all(funs(func_paste)) %>%
  rename("R1-1" = `1`,
         "R2-1" = `2`,
         "R1-2" = `3`,
         "R2-2" = `4`) %>%
  select(-type) %>%
  mutate_all(funs(as.character(.)))

no_of_reads <- as.data.frame(rbind(as.matrix(two_reads), as.matrix(four_reads))) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_hf150bp <- read.table("../data/mlst/hiseq_flex_150bp_mlst_summarized_results.tsv",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_hx125bp <- read.table("../data/mlst/hiseq_xt_125bp_mlst_summarized_results.tsv",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_hrx250bp <- read.table("../data/mlst/hiseqr_xt_250bp_mlst_summarized_results.tsv",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_data <- rbind(mlst_hf150bp, mlst_hx125bp, mlst_hrx250bp) %>%
  separate(ST.adk.fumC.gyrB.icd.mdh.purA.recA,
           into = c("x","ST","adk","fumC","gyrB","icd","mdh","purA","recA"),
           sep = " ") %>%
  select(-x) %>%
  rename("ref" = header) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)
```

# Sequence data
## Distribution of number of reads per sequencer
```{r}
sequence_data <- no_of_reads %>%
  left_join(isolate_data, by = "id") %>%
  mutate_at(vars(`R1-1`,`R2-1`,`R1-2`,`R2-2`),
            funs(as.numeric(as.character(.))))

ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
  geom_density(aes(fill = sequencer,
                   y = ..scaled..),
               alpha = 0.6)+
  labs(x = "Number of reads",
       y = "Density",
       fill = "Sequencer")+
  theme(legend.position = c(0.7, 0.8))
```

## Distribution of number of reads per library prep type
```{r}
ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
  geom_density(aes(fill = libprep,
                   y = ..scaled..),
               alpha = 0.6)+
  labs(x = "Number of reads",
       y = "Density",
       fill = "Library prep kit")+
  theme(legend.position = c(0.7, 0.8))
```

## Distribution of number of reads per animal species
```{r}
ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
  geom_density(aes(fill = species,
                   y = ..scaled..),
               alpha = 0.6)+
  labs(x = "Number of reads",
       y = "Density",
       fill = "Animal species")+
  theme(legend.position = c(0.7, 0.8))
```

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx","qep")
mut_genes <- c("gyr","par","marR","soxR","acrR","mcb")

# adjust parameters for filtering
if (length(ac_genes) == 1) {
  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
    ac_genes <- "ALL"
    } else {
      ac_genes <- ac_genes
    }
}

if (length(mut_genes) == 1) {  
  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
    mut_genes <- "ALL"
    } else {
      mut_genes <- mut_genes
    }
}
# -------------------------------------- Libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
               forcats,purrr,stringr,kableExtra,
               knitr,IRdisplay,reprex,svglite)

# -------------------------------------- Functions

func_paste2 <- function(x) paste(unique(x[!is.na(x)]), collapse = ",")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Identifies filenames in input folder
file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = "amr_report.tsv")
  return(files)
}

# Import ariba data from report.tsv from chosen database used in ariba
get_ariba_data <- function(filepath) {
  files <- file_names(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  names(data_list) <- files
  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
  return(data)
}

# Corrects the gene names found in the "cluster" column
fix_gene_names <- function(df) {
  genes <- unique(df$ref_name)
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  gene_names <- c()
  for (i in new_names) {
    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
    gene_names <- c(gene_names,p)
  }
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    rename(ref_name = genes)

  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

  return(df)
}

# Function for selecting genes of interest and filtering the 
# columns in the data frame on the resulting vector
select_genes <- function(df, gene_string) {
  names_df <- names(df)
  x <- "ref"
  for (i in names_df) {
    for (j in gene_string) {
      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
        x <- c(x, i)
      }
    }
  }
  df <- df %>%
    select(one_of(x))
  return(df)
}

# Allowed flags from the ARIBA report
flag_selection <- c("19","27","147","155","403",
                    "411","915","923","787","795",
                    "531","539","659","667","787","795")

# Function that returns info on flag selection
check_flags <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(flag_result = flag %in% flag_selection,
           flag_result = as.integer(flag_result)) %>%
    rename("gene" = gene_names)
  return(df)
}

# Function that handles megares data and returns a data frame with information on whether a gene
# is mutated or not. Includes control for QRDR in gyrA.
create_mut_table <- function(df) {
  df <- df %>%
    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
    select(ref, gene_names, flag, mutation) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, mutation) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut",
           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
  return(df)
}

# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
           gyrA_result = as.integer(gyrA_result),
           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           parC_result = mut %>% # control for mutation within QRDR for parC
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           parE_result = mut %>% # control for mutation within QRDR for parE
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    group_by(gene, result_total) %>%
    count() %>%
    ungroup() %>%
    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1),
           lwr = round(get_binCI(Present, Total)[1], 1),
           upr = round(get_binCI(Present, Total)[2], 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    gather(id, mut, -c(ref, gene, result, type))

  df2 <- fix_gyr_par_results(df1)

  df3 <- df2 %>%
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    spread(gene, mut) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

# creates a data frame with one row per sample, and 1/0 results 
# for mutations in respective genes in columns
create_mut_report <- function(df) {
  df <- df %>%
    select(-mut) %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# creates a data frame with one row per sample, and 1/0 results
# for acquired genes in columns
create_acquired_report <- function(df) {
  df <- df %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# -------------------------------------- Analysis
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

# Clean data
clean_mut_data <- fix_gene_names(mut_data)
clean_acquired_data <- fix_gene_names(acquired_data)

# Check flags
mut_flags <- check_flags(clean_mut_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

acquired_flags <- check_flags(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

# Wrangle
mut_table <- create_mut_table(clean_mut_data)
mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

mut_filtered <- filter_mut_table(mut_table_fixed)
acquired_table <- create_acquired_table(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

acquired_filtered <- filter_acquired_table(acquired_table)

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

# ---------------------------------- Tables

# number of mutations table
mut_no <- mut_quant %>%
  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
  count()

# number of isolates per mutation combination
mut_comb <- mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  count()
```

# Tables
## Isolate data
```{r}
isolate_data_complete <- isolate_data %>%
  left_join(no_of_reads, by = "id")

isolate_data_complete
```

## List of genes of interest
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

acquired_ref_genes <- acquired_data %>%
  mutate(test = duplicated(ref_name),
         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9")) %>%
  filter(test == FALSE,
         ref_name %in% acquired_ref_names) %>%
  select(ref_name, gene) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
         Type = "Acquired")

df_ref_names <- mut_data$cluster
ref_names <- c()

for (i in mut_genes) {
  for (j in df_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      ref_names <- c(ref_names, j)
    }
  }
}

ref_names <- unique(ref_names)

intrinsic_ref_genes <- mut_data %>%
  mutate(test = duplicated(cluster),
         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
                          grepl("soxs", cluster, ignore.case = T) == TRUE ~ "soxS",
                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR")) %>%
  filter(test == FALSE,
         cluster %in% ref_names) %>%
  select(ref_name, gene, free_text) %>%
  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
  select(-ref_name) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
                                 Gene == "soxS" ~ "Transcriptional activator of the superoxide response regulon.",
                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump."),
         Type = "Intrinsic gene")

ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)

ref_genes

```

## Isolates per species
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()

per_species
```

## Percent isolates with PMQR
```{r}
acquired_report %>%
  select(-id) %>%
  mutate_all(funs(as.numeric(.))) %>%
  mutate(total = rowSums(.),
         result = if_else(total != 0, 1, 0)) %>%
  group_by(result) %>%
  count() %>%
  ungroup() %>%
  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
  spread(result, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,2),
         lwr = round(get_binCI(Present, Total)[1], 2),
         upr = round(get_binCI(Present, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr))
```

## Percent isolates with specific qnr genes
```{r}
acquired_report %>%
  select(-id) %>%
  t() %>%
  as.data.frame(.) %>%
  mutate(Gene = rownames(.)) %>%
  mutate_at(vars(-Gene),
            funs(as.numeric(as.character(.)))) %>%
  mutate(Present = rowSums(.[,-286])) %>%
  select(c(Gene, Present)) %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(Present/Total * 100,2),
         lwr = round(get_binCI(Present, Total)[1], 2),
         upr = round(get_binCI(Present, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Percent isolates with mutations in QRDR
```{r}
mut_report %>%
  select(-c(id, marR, nt_change)) %>%
  group_by_all() %>%
  count() %>%
  ungroup() %>%
  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
            .funs = funs(as.numeric)) %>%
  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
  select(n, type) %>%
  group_by(type) %>%
  summarise_all(funs(sum)) %>%
  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
  spread(type, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,2),
         lwr = round(get_binCI(Present, Total)[1], 2),
         upr = round(get_binCI(Present, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Percent of isolates with specific mutation combinations in QRDR
```{r}
mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  count() %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(n/Total * 100,2),
         lwr = round(get_binCI(n, Total)[1], 2),
         upr = round(get_binCI(n, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Percent isolates with specific nucleotide changes in QRDR
```{r}
suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  select(-c(id, flag)) %>%
  gather(gene, mut, -ref) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  select(-mut) %>%
  rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  group_by(gene, mut, nt_change) %>%
  count() %>%
  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(n/Total * 100,1),
         lwr = round(get_binCI(n, Total)[1],1),
         upr = round(get_binCI(n, Total)[2],1),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total)) %>%
  rename("Mutation" = mut))
```

## Total combinations
```{r}
test <- acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("marR","id")], by = "id") %>%
  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  select(-id) %>%
  group_by_at(vars(-c(CIP, NAL))) %>%
  summarise(CIP = ifelse(min(as.numeric(CIP)) == max(as.numeric(CIP)),
                         as.character(median(as.numeric(CIP))),
                         paste(min(as.numeric(CIP)), max(as.numeric(CIP)), sep = " - ")),
            NAL = ifelse(min(as.numeric(NAL)) == max(as.numeric(NAL)),
                         as.character(median(as.numeric(NAL))),
                         paste(min(as.numeric(NAL)), max(as.numeric(NAL)), sep = " - ")),
            n = n()) %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(n/Total * 100,1),
         lwr = round(get_binCI(n, Total)[1],1),
         upr = round(get_binCI(n, Total)[2],1),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Significant differences between resistance mechanisms for each animal species
```{r, warning = F, message = F}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

do_chisq_test <- function(df) {
  df <- df %>%
    slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
          filter(Var2 < Var1) %>% t) %>%
    mutate(test = rep(1:(n()/2), each = 2)) %>%
    group_by(test) %>%
    do(data_frame(gene = .$key,
                  test = first(.$test),
                  species1 = first(.$species),
                  species2 = last(.$species),
                  data = list(matrix(c(.$Absent,
                                       .$Present),
                                     ncol = 2)))) %>%
    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
    mutate(p.value = map_dbl(chi_test, "p.value"),
           dupl = duplicated(test)) %>%
    filter(dupl == FALSE) %>%
    select(gene, species1, species2, p.value) %>%
    filter(p.value < 0.05)
  return(df)
}

total_df <- acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("marR","id")], by = "id") %>%
  left_join(isolate_data[c("id", "species")])

col_proc <- c("gyrA","gyrB","parC","parE")

id <- total_df$id

cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  cbind(id) %>%
  gather(key, value, -id) %>%
  filter(grepl("(.*?)_0", key) == FALSE) %>%
  spread(key, value)

mechanism_per_species <- total_df %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(cols_to_add, by = "id") %>%
  select(id, species, everything()) %>%
  select(-id) %>%
  gather(key, value, -species) %>%
  group_by_all() %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  select(key, species, Present, Absent) %>%
  split(f = .$key)

p_values_total <- suppressWarnings(lapply(mechanism_per_species, function(x) do_chisq_test(x))) %>%
  bind_rows()

p_values_total
```

# Figures
## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
```{r}
total_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  select(-ref) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(present,CIP) %>%
  count() %>%
  ungroup() %>%
  rowwise() %>%
  mutate(present = ifelse(present == "", "None", present),
         total = 285,
         percent = round(n/total * 100, 1),
         lwr = round(get_binCI(n, total)[1], 1),
         upr = round(get_binCI(n, total)[2], 1))

genes <- c(names(mut_report), names(acquired_report))
genes <- genes[!genes == "id"]

p1 <- ggplot(total_report, aes(reorder(present, as.numeric(-CIP)), percent, fill = factor(CIP)))+
  geom_col(color = "black")+
  labs(y = "Percent (%) isolates")+
  scale_fill_viridis(discrete = TRUE)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top",
        legend.spacing.x = unit(0.27, "cm"))

dot_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,CIP)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit()

p2 <- ggplot(dot_report, aes(reorder(present, -CIP), gene, fill = gene))+
  geom_point(size = 2, pch = 21)+
  scale_fill_viridis(option = "B",
                     discrete = TRUE)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        axis.ticks.x = element_blank())

plot_grid(p1, p2, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))

```

## Percent QREC isolates with mutations in intrinsic genes
```{r}
gene_names <- names(mut_report)

gene_names <- gene_names[-c(1,2)]

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

## Percent QREC isolates with Qnr genes per species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

acquired_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(0.9)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
    )
```

## Total resistance mechanisms per animal species
```{r, warning=F, message=F, fig.height=8, fig.width=12}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

total_df <- acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("marR","id")], by = "id") %>%
  left_join(isolate_data[c("id", "species")])

col_proc <- c("gyrA","gyrB","parC","parE")

id <- total_df$id

cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  cbind(id) %>%
  gather(key, value, -id) %>%
  filter(grepl("(.*?)_0", key) == FALSE) %>%
  spread(key, value)

total_df2 <- total_df %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(cols_to_add, by = "id") %>%
  select(id, species, everything()) %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,4:10]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-id) %>%
  mutate_at(vars(-species, CIP),
            funs(as.numeric(as.character(.)))) %>%
  group_by(species, CIP) %>%
  summarise_all(funs(sum, n = n())) %>%
  select(-c(qepA4_n,qnr_n,gyrA_D87G_n,gyrA_D87H_n,
            gyrA_D87N_n,gyrA_D87Y_n,gyrA_S83A_n,
            gyrA_S83L_n,parC_A56T_n,parC_E84V_n,
            parC_S57T_n,parC_S58I_n,parC_S80I_n,
            parC_S80R_n,parE_A512T_n,parE_D463N_n,
            parE_D475E_n,parE_H516Y_n,parE_L416F_n,
            parE_L488M_n,parE_S458A_n)) %>%
  gather(gene, value, -c(species, CIP, marR_n)) %>%
  rename("n" = marR_n) %>%
  mutate(gene2 = sub("(.*?)_(.*?)_sum", "\\1", gene),
         gene2 = sub("(.*?)_sum", "\\1", gene2),
         gene = sub("(.*?)_(.*?)_sum", "\\2", gene),
         gene = sub("(.*?)_sum", "\\1", gene),
         type = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "qnr" ~ 6,
                          gene2 == "qepA4" ~ 7),
         perc = value/n*100)
# 
# ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
#   geom_tile(color = "grey70")+
#   scale_fill_viridis(option = "A",
#                      begin = 0.1,
#                      guide = guide_colorbar(ticks = F,
#                                             barheight = 0.5,
#                                             barwidth = 10))+
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 7),
#         axis.text.y = element_text(size = 8),
#         axis.title.x = element_blank(),
#         axis.title.y = element_text(size = 10),
#         legend.title = element_blank(),
#         legend.position="bottom",
#         legend.justification="center",
#         legend.direction="horizontal",
#         legend.text = element_text(size = 8),
#         strip.background = element_rect(colour = "black", fill = "grey70"),
#         strip.text = element_text(size = 10))+
#   labs(y = "CIP MIC")+
#   facet_wrap(~species)+
#   coord_fixed()

p1 <- ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
  geom_tile(color = "grey70")+
  scale_fill_viridis(option = "A",
                     begin = 0.1)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        strip.text = element_text(size = 10),
        plot.margin = unit(c(0,0,0,0), "cm"))+
  labs(y = "CIP MIC")+
  facet_wrap(~species, ncol = 1)+
  coord_fixed()

nplot_data <- total_df %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(species, CIP) %>%
  count() %>%
  ungroup() %>%
  mutate(species = paste0(species, " n"))

p2 <- ggplot(nplot_data, aes(factor(CIP), n))+
  geom_col()+
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 10),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        plot.margin = unit(c(0,0,-10,0), "cm"))+
  facet_wrap(~species, ncol = 1)+
  coord_flip()

p1g <- ggplotGrob(p1)
p2g <- ggplotGrob(p2)

g <- cbind(p1g, p2g, size = "first")

grid.arrange(g)

```