### 2019-02-07 notebooks/assembly_metrics.Rmd
added boxplots on summary metrics to verify similarity between isolates from animal species and sequencer

---
title: "Assembly Quality Metrics - Quast Results"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r, include = FALSE}
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, scales)

"%not_in%" <- Negate("%in%")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

```

```{r}
path <- "../data/quast/"

file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = ".tsv")
  return(files)
}

create_df <- function(df) {
  df <- df %>%
    dplyr::rename(ctg_0_bp = X..contigs.....0.bp.,
           ctg_1000_bp = X..contigs.....1000.bp.,
           ctg_5000_bp = X..contigs.....5000.bp.,
           ctg_10000_bp = X..contigs.....10000.bp.,
           ctg_25000_bp = X..contigs.....25000.bp.,
           ctg_50000_bp = X..contigs.....50000.bp.,
           tot_len_0_bp = Total.length.....0.bp.,
           tot_len_1000_bp = Total.length.....1000.bp.,
           tot_len_5000_bp = Total.length.....5000.bp.,
           tot_len_10000_bp = Total.length.....10000.bp.,
           tot_len_25000_bp = Total.length.....25000.bp.,
           tot_len_50000_bp = Total.length.....50000.bp.,
           no_of_contigs = X..contigs,
           no_of_misassemblies = X..misassemblies,
           no_of_misassembled_ctg = X..misassembled.contigs,
           no_of_local_misassemblies = X..local.misassemblies,
           unaligned_contigs = X..unaligned.contigs,
           N_per_100kbp = X..N.s.per.100.kbp,
           mismatches_per_100kbp = X..mismatches.per.100.kbp,
           indels_per_100kbp = X..indels.per.100.kbp,
           genes = X..genes,
           GC = GC....,
           duplication_ratio = Duplication.ratio,
           largest_contig = Largest.contig)
  return(df)
}

```

# Import files and clean data
```{r}
files <- file_names(path)

data_list <- lapply(files,
                    FUN = function(file) {
                      read.delim(paste0(path, "/", file),
                                 sep = "\t",
                                 header = TRUE,
                                 stringsAsFactors = FALSE)
                    })

data_list <- lapply(data_list, function(df) mutate_all(df, as.character))

names(data_list) <- files

df <- bind_rows(data_list)

df_clean <- create_df(df) %>%
  filter(grepl("broken", Assembly) == FALSE) %>%
  filter(Assembly %not_in% c("51-2016-02-486_S31_pilon_spades",
                             "68-2016-17-565_S45_pilon_spades",
                             "52-2016-02-732_S43_pilon_spades",
                             "50-2016-02-428_S19_pilon_spades",
                             "4-2014-01-1741_TAAGGCGA-ACTGCATA_pilon_spades")) %>%
  mutate(ref = Assembly %>%
           gsub("_pilon_spades", "", .)) %>%
  left_join(id_data, by = "ref") %>%
  left_join(isolate_data[c("id","sequencer")], by = "id")

metrics <- df_clean %>%
  select(id, no_of_contigs, largest_contig, GC, N50, NG50, L50, LG50, no_of_misassemblies, duplication_ratio, genes) %>%
  mutate_at(vars(no_of_contigs, largest_contig, N50),
            funs(as.numeric))

metrics

metrics2 <- metrics %>%
  left_join(isolate_data, by = "id")

embed_data <- function(x = mtcars, filename = "file.csv", label= "Get data") {

  # Create encoded Base64 datastream 
  encode_data <- function(x){
    write.csv2(x, "./file.csv")
    enc= sprintf('data:text/csv;base64,%s', openssl::base64_encode(paste0(readLines("./file.csv"), collapse="\n")) )
    unlink("./file.csv")
    return(enc)
  }

  # String result ready to be placed in rmarkdown
  paste0("<a download='", filename, "' href=", encode_data(x), ">", label, "</a>")

}
```

`r embed_data(metrics, filename = "assembly_metrics.csv")`

# Mean metrics
```{r}
print(paste0("Mean number of contigs: ", round(mean(metrics2$no_of_contigs), 1)))
print(paste0("Mean size of largest contigs: ", round(mean(metrics2$largest_contig), 1)))
print(paste0("Mean N50: ", round(mean(metrics2$N50), 1)))
print(paste0("Mean NG50: ", round(mean(as.numeric(metrics2$NG50)), 1)))

```

# Summarised assembly stats
## Summary stats per animal species
```{r}
p1 <- ggplot(metrics2, aes(species, N50)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

p2 <- ggplot(metrics2, aes(species, no_of_contigs)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  labs(y = "No. of contigs") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

p3 <- ggplot(metrics2, aes(species, largest_contig)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  labs(y = "Largest contig") +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

p4 <- ggplot(metrics2, aes(species, as.numeric(no_of_misassemblies))) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  labs(y = "No. of misassemblies") +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

plot_grid(p1, p2, p3, p4, ncol = 2, nrow = 2, align = "v")
```

## Summary stats per sequencing method
```{r}
p5 <- ggplot(metrics2, aes(sequencer, N50)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

p6 <- ggplot(metrics2, aes(sequencer, no_of_contigs)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  labs(y = "No. of contigs") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

p7 <- ggplot(metrics2, aes(sequencer, largest_contig)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  labs(y = "Largest contig") +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

p8 <- ggplot(metrics2, aes(sequencer, as.numeric(no_of_misassemblies))) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot() +
  labs(y = "No. of misassemblies") +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8))

plot_grid(p5, p6, p7, p8, ncol = 2, nrow = 2, align = "v")
```