### 2019-03-27 notebooks/article_notebook2.Rmd
minor changes

---
title: "Article data analysis 2"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(viridis)
library(purrr)
library(tibble)
library(stringr)
library(Biostrings)
library(seqinr)
library(scales)
library(gridExtra)
```

```{r Import data, include = FALSE}
# Vector of excluded samples
ex_samples <- c(
  "2016-17-565-1-S",
  "2016-02-428-2-S",
  "2016-02-486-2-S",
  "2016-02-732-2-S",
  "2014-01-1741-1-S"
  )

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

"%not_in%" <- Negate("%in%")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

two_reads <- read.table("../data/no_of_reads.txt",
                          sep = "\t",
                          header = F,
                          stringsAsFactors = F) %>%
  dplyr::rename("ref" = V1,
         "reads" = V2) %>%
  mutate(ref = gsub(".+/(.*?)\\.fastq\\.gz", "\\1", ref),
         type = ifelse(grepl("R1", ref) == TRUE, "R1", "R2")) %>%
  mutate(ref = gsub("(.*?)_L00[0-9]_R[1-2]_00[0-9]","\\1", ref)) %>%
  group_by(ref) %>%
  mutate(id = n()) %>%
  {. ->> four_reads} %>%
  filter(id == 2) %>%
  spread(type, reads) %>%
  dplyr::rename("R1-1" = R1,
         "R2-1" = R2) %>%
  mutate("R1-2" = NA,
         "R2-2" = NA) %>%
  select(-id) %>%
  mutate_all(funs(as.character(.)))

four_reads <- four_reads %>%
  filter(id == 4) %>%
  group_by(ref) %>%
  mutate(id = 1:n()) %>%
  spread(id, reads) %>%
  summarise_all(funs(func_paste)) %>%
  dplyr::rename("R1-1" = `1`,
         "R2-1" = `2`,
         "R1-2" = `3`,
         "R2-2" = `4`) %>%
  select(-type) %>%
  mutate_all(funs(as.character(.)))

no_of_reads <- as.data.frame(rbind(as.matrix(two_reads),
                                   as.matrix(four_reads))) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

mlst_data <- read.table("../data/mlst_results.txt",
                        sep = "\t",
                        header = TRUE,
                        stringsAsFactors = F) %>%
  filter(id %not_in% "2014-01-1741-1-S")
```

# ARIBA reports analyses
This script filters results from ARIBA reports for genes of interest
```{r ARIBA analysis}
# Report directories
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"

# Genes of interest. These genes are partially matched to 
# the generated gene names from the "fix_gene_names" function
ac_genes <- c("qnr","oqx","qep","aac6")
mut_genes <- c("gyr","par","marR","marA",
               "sox","rpoB","rob")

# -------------------------- Functions ----------------------------

# Paste unique cell values together, separate with ","
func_paste2 <- function(x) paste(unique(x[!is.na(x)]),
                                 collapse = ",")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(
  setNames(
    binom.test(x,n)$conf.int*100, 
    c("lwr", "upr")
    )
  )

# Identifies filenames in input folder
file_names <- function(filepath) {
  files <- list.files(path = filepath, 
                      pattern = "amr_report.tsv")
  return(files)
}

# Import ariba data from report.tsv identified with file_names
get_ariba_data <- function(filepath) {
  # Identify file names in filepath
  files <- file_names(filepath)

  # Import data
  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  # Set file names in list
  names(data_list) <- files

  # Bind to data frame and convert columns to character
  data <- bind_rows(
    lapply(
      data_list,
      function(x) map(x, as.character)
      ),
    .id = "ref")

  return(data)
}

# Corrects gene names 
fix_gene_names <- function(df) {
  # Identify unique entries in ref_name column
  genes <- unique(df$ref_name)

  # Strip away excess characters
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  # make all genes lower case, with last letter 
  # upper case if four characters long
  gene_names <- c()

  for (i in new_names) {
    p <- paste(
      tolower(
        substring(i, 1,3)
        ),
      substring(i, 4),
      sep = "",
      collapse = " ")
    gene_names <- c(gene_names,p)
  }

  # Create data frame for matching names below
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    dplyr::rename(ref_name = genes)

  # Left join to match with ref_name column
  # and strip ref column for file suffix
  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref)) %>%
    mutate(gene_names = ifelse(gene_names == "soxS", "soxR", gene_names))

  return(df)
}

# Function for selecting genes of interest and filtering the 
# columns in the data frame on the resulting vector
select_genes <- function(df, gene_string) {
  names_df <- names(df)
  x <- "ref"
  for (i in names_df) {
    for (j in gene_string) {
      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
        x <- c(x, i)
      }
    }
  }
  df <- df %>%
    select(one_of(x))
  return(df)
}

# Allowed flags from the ARIBA report
flag_selection <- c("19","27","147","155","403",
                    "411","915","923","787","795",
                    "531","539","659","667","787","795")

# Function that returns info on flag selection
check_flags <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change, ref_nt, ctg_nt) %>%
    mutate(flag_result = flag %in% flag_selection,
           flag_result = as.integer(flag_result),
           nt_change = paste(ref_nt, ctg_nt, sep = "-")) %>%
    mutate(nt_change = ifelse(nt_change == ".-.", "0", nt_change)) %>%
    dplyr::rename("gene" = gene_names) %>%
    select(-c(ref_nt, ctg_nt))
  return(df)
}

# Function that handles megares data and returns a data 
# frame with information on whether a gene is mutated 
# or not
create_mut_table <- function(df) {
  df <- df %>%
    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
    select(ref, gene_names, flag, mutation) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, mutation) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut",
           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
  return(df)
}

# Corrects for mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    # control for mutation within QRDR for gyrA
    mutate(gyrA_result = mut %>% 
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             # returns TRUE/FALSE whether value is within range or not
             map_lgl(~ any(.x >= 67L & .x <= 106L)), 
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result),
           gyrA_result = as.integer(gyrA_result),
           # control for mutation within QRDR for gyrB
           gyrB_result = mut %>%
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           # control for mutation within QRDR for parC
           parC_result = mut %>% 
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           # control for mutation within QRDR for parE
           parE_result = mut %>% 
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    # mergeresults from all four genes into one column
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    # merge results from other genes with the four genes
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"),
                                 result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    # Filter flags
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based 
# on the vector "mut_genes"
filter_mut_table <- function(df) {
  # Identify unique genes in column
  report_genes <- unique(df$gene)
  grep_genes <- c()

  # match names from name vector to column entries
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }

  # Filter data frame based on values in grep_genes
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based
# on the string "ac_genes"
filter_acquired_table <- function(df) {
  # Identify unique genes in column
  report_genes <- unique(df$gene)
  grep_genes <- c()

  # match names from vector to column entries
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }

  # Filter data frame based on values in grep_genes
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    # Count Presence/Absence of each gene
    group_by(gene, result_total) %>%
    dplyr::count() %>%
    ungroup() %>%
    # Convert binary to "Present" and "Absent"
    mutate(result_total = if_else(result_total == 1, 
                                  "Present", "Absent")) %>%
    # Spread to separate Present and Absent columns
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    # Calculate percentages
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    # Filter out all other genes
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    select(-nt_change) %>%
    # Count how many mutations in each gene per row
    mutate(entries = sapply(strsplit(.$mut, ","),
                            FUN = function(x) {length(x)})) %>%
    # Separate each mutation to own column
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    # Gather all mutations together into one column,
    # to get one row per mutation
    gather(id, mut, -c(ref, gene, result, type))

  # Correct for QRDR mutations in the four genes
  df2 <- fix_gyr_par_results(df1)

  # Generate final table
  df3 <- df2 %>%
    # Filter out mutations outide QRDR
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    # Create own column for each gene
    spread(gene, mut) %>%
    group_by(ref) %>%
    # Summarise all columns based on unique values
    summarise_all(funs(func_paste)) %>%
    # Remove "0" values for correct counts below
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    # Count mutations in each column entry
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","),
                             FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","),
                             FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","),
                             FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","),
                             FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    # Reinsert "0" values 
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

# creates a data frame with one row per sample, and 1/0 results 
# for mutations in respective genes in columns
create_mut_report <- function(df) {
  df <- df %>%
    select(-mut) %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# creates a data frame with one row per sample, and 1/0 results
# for acquired genes in columns
create_acquired_report <- function(df) {
  df <- df %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# -------------------------- Analysis ---------------------------
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

# Clean data
clean_mut_data <- fix_gene_names(mut_data)
clean_acquired_data <- fix_gene_names(acquired_data)

# Check flags
mut_flags <- check_flags(clean_mut_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  # Filters out unwanted samples
  filter(id %not_in% ex_samples)

acquired_flags <- check_flags(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

mut_table <- create_mut_table(clean_mut_data)

mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

mut_filtered <- filter_mut_table(mut_table_fixed)

acquired_table <- create_acquired_table(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

acquired_filtered <- filter_acquired_table(acquired_table)

mut_stats <- calc_stats(mut_filtered)
mut_report <- create_mut_report(mut_filtered)

acquired_stats <- calc_stats(acquired_filtered)
acquired_report <- create_acquired_report(acquired_filtered)

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  # Correct for erroneous entry from ARIBA
  mutate(parC = ifelse(parC == "S58I", "S80I", parC)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

# number of mutations table
mut_no <- mut_quant %>%
  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
  dplyr::count()

# number of isolates per mutation combination
mut_comb <- mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  dplyr::count()
```

# Tables
## Isolate data
This table presents the isolates and their accompanying metadata.
```{r}
isolate_data_complete <- isolate_data %>%
  # combine with read data
  left_join(no_of_reads, by = "id")

isolate_data_complete
```

## Isolates per species
This table presents how many isolates were included per species.
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  dplyr::count()

per_species
```

## Sequence types in total
```{r}
tot_mlst <- mlst_data[,c("id","ST")] %>%
  group_by(ST) %>%
  dplyr::count() %>%
  arrange(desc(n))

print(paste0("Total amount of different sequence types: ", nrow(tot_mlst)))

tot_mlst
```

## Sequence types per animal species
This table presents all the different sequence types identified in each animal species, arranged in a descending order.
```{r}
mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(ST, species) %>%
  dplyr::count() %>%
  arrange(desc(n))
```

## Most abundant sequence types per animal species
This table presents the most abundant sequence types identified in each animal species (n > 3).
```{r}
# Identify which STs have more than three observations
mlst_n <- mlst_data[,c("id","ST")] %>%
  group_by(ST) %>%
  dplyr::count() %>%
  # filter out all STs with less than four observations
  filter(n > 3)

mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  spread(ST, species) %>%
  summarise_all(funs(func_paste)) %>%
  select(-id) %>%
  gather(ST, species) %>%
  # filter data frame on the STs identified above
  filter(ST %in% mlst_n$ST)
```

## Number of sequence types per animal species
This table presents the amount of different sequence types that were identified in each animal species in total.
```{r}
mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(species) %>%
  # count how many STs per group
  mutate(n = length(unique(ST))) %>%
  select(species,n) %>%
  summarise_all(funs(func_paste))
```

## Percent isolates with PMQR
This table presents the percentage of isolates with identified PMQR determinants in total.
```{r}
acquired_report %>%
  select(-id) %>%
  mutate_all(funs(as.numeric(.))) %>%
  # presence/absence of qnr genes in total
  mutate(total = rowSums(.),
         result = if_else(total != 0, 1, 0)) %>%
  group_by(result) %>%
  # count groups
  dplyr::count() %>%
  ungroup() %>%
  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
  spread(result, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1))
```

## Percent isolates with qnr family genes
```{r}
acquired_report %>%
  select(-c(qepA4, id)) %>%
  mutate_all(funs(as.numeric)) %>%
  dplyr::rename("qnrA" = qnrA1) %>%
  # identify presence/absence for family groups 
  # from the gene subgroups
  mutate(qnrB = ifelse(rowSums(.[,2:4]) >= 1, 1, 0),
         qnrS = ifelse(rowSums(.[,5:7]) >= 1, 1, 0)) %>%
  select(qnrA, qnrB, qnrS) %>%
  gather(gene, value) %>%
  group_by(gene, value) %>%
  # count groups
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 0, "Absent", "Present")) %>%
  spread(value, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1))
```

## Percent isolates with specific PMQR genes
```{r}
acquired_report %>%
  select(-id) %>%
  # transpose table
  t() %>%
  as.data.frame(.) %>%
  mutate(Gene = rownames(.)) %>%
  mutate_at(vars(-Gene),
            funs(as.numeric(as.character(.)))) %>%
  mutate(Present = rowSums(.[,-281])) %>%
  select(c(Gene, Present)) %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(Present/Total * 100,1))
```

## qnr-genes per animal species
This table presents the specific qnr-genes identified in each animal species.
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")

acquired_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, qnr_genes) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  arrange(key, desc(Percent))
```

## Percent isolates with mutations in QRDR
This table presents the percent of isolates with mutations in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
```{r}
mut_report %>%
  select(gyrA, gyrB, parC, parE) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
            .funs = funs(as.numeric)) %>%
  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
  select(n, type) %>%
  group_by(type) %>%
  summarise_all(funs(sum)) %>%
  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
  spread(type, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1)) %>%
  select(-Total)
```

## Percent mutations in each intrinsic gene
This table presents the percent mutations in all intrinsic _E. coli_ genes included in this study.
```{r}
mut_report %>%
  select(-nt_change) %>%
  gather(gene, value, -id) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1)) %>%
  arrange(desc(Percent))
```

## Percent mutations in each intrinsic gene per animal species
```{r}
mut_report %>%
  select(-nt_change) %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  gather(gene, value, -c(id, species)) %>%
  group_by(gene, species, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1)) %>%
  arrange(gene)
```

## Specific mutations in QRDR
This table presents the specific mutations identified in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
```{r, warning=FALSE, message=FALSE}
clean_mut_data %>%
  # correct non-mutated entries
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".",
                                 "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  # filter quality flags and genes
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  # create columns for each gene
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  # summarise for each sample
  summarise_all(funs(func_paste2)) %>%
  select(-c(id, flag)) %>%
  # set correct sample id's
  left_join(id_data, by = "ref") %>%
  dplyr::select(id, everything(), -ref) %>%
  # filter out excluded samples
  filter(id %not_in% ex_samples) %>%
  gather(gene, mut, -id) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE,
                      0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  # create one row per mutation
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  # correct erroneous result from ARIBA
  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I",
                            "S80I", aa_change)) %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  # fix results for the four genes
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  group_by(gene, mut, nt_change) %>%
  # count groups
  dplyr::count() %>%
  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1)) %>%
  dplyr::rename("Mutation" = mut) %>%
  arrange(gene, desc(Percent))
```

## Specific QRDR mutations per animal species
```{r, warning=FALSE, message=FALSE}
clean_mut_data %>%
  # correct non-mutated entries
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".",
                                 "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  # filter quality flags and genes
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  # create columns for each gene
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  dplyr::select(-c(id, flag)) %>%
  left_join(id_data, by = "ref") %>%
  dplyr::select(id, everything(), -ref) %>%
  # filter out excluded samples
  filter(id %not_in% ex_samples) %>%
  gather(gene, mut, -id) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE,
                      0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  # correct errouneous result from ARIBA
  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I",
                            "S80I", aa_change)) %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  # fix results for the four genes
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  group_by(gene, mut, species) %>%
  dplyr::count() %>%
  rowwise() %>%
  # set total amount of isolates per animal species
  mutate(Total = case_when(species == "Broiler" ~ 87,
                           species == "Pig" ~ 75,
                           species == "Red fox" ~ 52,
                           species == "Wild bird" ~ 66),
         Percent = round(n/Total * 100,1)) %>%
  dplyr::rename("Mutation" = mut) %>%
  arrange(gene, desc(Percent))
```

## Mutation combinations in QRDR
This table presents all the observed combinations of mutations in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
```{r}
mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  # count combinations across columns
  dplyr::count() %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1)) %>%
  arrange(desc(Percent))
```

## Mutations in intrinsic genes per animal species
This table presents the percentage of isolates with mutations for each intrinsic gene.
```{r}
# Identify gene names
gene_names <- names(mut_report)
gene_names <- gene_names[-c(1,2)]

mut_report %>%
  left_join(isolate_data, by = "id") %>%
  # use gene names to specify which columns to gather
  gather(key, value, gene_names) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  dplyr::rename("Gene" = key,
                "Species" = species) %>%
  dplyr::select(Gene, Species, Present, Total, Percent) %>%
  arrange(Gene, desc(Percent))
```

## Specific mutations in non-QRDR intrinsic genes
These mutations are the most prominent in each gene where mutations were identified (n > 2). The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
```{r, message=FALSE, warning=FALSE}
mut_filtered %>%
  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
  select(id, gene, mut) %>%
  mutate(entries = sapply(strsplit(.$mut, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
  mutate(value = str_trim(value)) %>%
  select(id, gene, value) %>%
  filter(value != 0,
         is.na(value) == FALSE) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  filter(n > 2) %>%
  mutate(Total = case_when(
    gene == "marA" ~
      length(which(mut_report[, "marA"] == 1)),
    gene == "marR" ~
      length(which(mut_report[, "marR"] == 1)),
    gene == "rpoB" ~
      length(which(mut_report[, "rpoB"] == 1)),
    gene == "soxR" ~
      length(which(mut_report[, "soxR"] == 1))),
         percent = case_when(
    gene == "marA" ~ 
      round(n / Total*100, 1),
    gene == "marR" ~ 
      round(n / Total*100, 1),
    gene == "rpoB" ~ 
      round(n / Total*100, 1),
    gene == "soxR" ~ 
      round(n / Total*100, 1))) %>%
  arrange(gene, desc(percent)) %>%
  dplyr::rename("Gene" = gene,
                "Substitution" = value,
                "Percent" = percent)

```

## Mechanisms in isolates without QRDR mutations
This table presents the mechanisms identified in isolates with no mutations in either of _gyrA_, _gyrB_, _parC_ and/or _parE_.
```{r}
specific_ac_genes <- acquired_report %>%
  gather(PMQR, value, -id) %>%
  filter(value == 1) %>%
  select(-value)

mut_report %>%
  # all gyrB were 0
  filter(gyrA == 0,
         parC == 0,
         parE == 0) %>%
  left_join(specific_ac_genes, by = "id") %>%
  select(-c(nt_change, id)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  dplyr::select(-c(gyrA,gyrB,parC,parE))
```

## Total combinations of resistance mechanisms
This table presents all combinations of resistance mechanisms included in this study, and their respective MIC.
```{r}
mut_report %>%
  select(-c(nt_change, gyrA, parC, parE)) %>%
  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  # summarise qnr-genes
  mutate(qnr = ifelse(rowSums(.[,12:18]) == 0, 0, 1)) %>%
  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  arrange(desc(CIP))
```

## Significant differences in resistance mechanisms
This table presents the significant (p < 0.05) differences in resistance mechanisms between species, identified with chi-squared tests.
```{r, message=FALSE, warning=FALSE}
# function for doing chi-squared tests on multiple entries in a list
do_chisq_test <- function(df) {
  df <- df %>%
    # create columns with all possible comparisons
    dplyr::slice(expand.grid(1:length(unique(species)),
                             1:length(unique(species))) %>%
                   rev %>% 
          dplyr::filter(Var2 < Var1) %>% t) %>%
    # create test ID
    mutate(test = rep(1:(n()/2), each = 2)) %>%
    group_by(test) %>%
    # specify values for chi-squred tests
    dplyr::do(data_frame(gene = .$key,
                  test = dplyr::first(.$test),
                  species1 = dplyr::first(.$species),
                  species2 = dplyr::last(.$species),
                  data = list(matrix(c(.$Absent,
                                       .$Present),
                                     ncol = 2)))) %>%
    # do chi-squared tests
    mutate(chi_test = map(data, chisq.test,
                          correct = FALSE)) %>%
    mutate(p.value = map_dbl(chi_test, "p.value"),
           dupl = duplicated(test)) %>%
    # filter out duplicated tests
    dplyr::filter(dupl == FALSE) %>%
    ungroup() %>%
    select(gene, species1, species2, p.value) %>%
    # round off p-value column
    mutate(p.value = round(p.value, 5)) %>%
    dplyr::filter(p.value <= 0.05)
  return(df)
}

# create total mechanism table
total_df <- acquired_report %>%
  left_join(mut_report, by = "id") %>%
  left_join(isolate_data[c("id", "species")], by = "id") %>%
  select(-nt_change)

# create list of genes with total presence/absence values
# from each animal species
mechanism_per_species <- total_df %>%
  select(id, species, everything()) %>%
  select(-id) %>%
  gather(key, value, -species) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  select(key, species, Present, Absent) %>%
  split(f = .$key)

# Get intrinsic gene names
intrinsic_names <- names(mut_report)
intrinsic_names <- intrinsic_names[-c(1,2)]

# create percentage table over intrinsic genes
intrinsic_gene_perc <- mut_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, intrinsic_names) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  select(key, species, Percent) %>%
  dplyr::rename("gene" = key)

# Get acquired gene names
acquired_names <- names(acquired_report)
acquired_names <- acquired_names[-1]

# create percentage table over acquired genes
acquired_genes_perc <- acquired_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, acquired_names) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  select(key, species, Percent) %>%
  dplyr::rename("gene" = key)

# create total percentage table
total_perc <- rbind(intrinsic_gene_perc, acquired_genes_perc)

# run chi-squared tests
p_values_total <- lapply(mechanism_per_species, function(x) do_chisq_test(x)) %>%
  # bind results
  bind_rows() %>%
  # specify percentages for species 1
  left_join(total_perc, by = c("gene", c("species1" = "species"))) %>%
  dplyr::rename("Percent species 1" = Percent) %>%
  # specify percentages for species 2
  left_join(total_perc, by = c("gene", c("species2" = "species"))) %>%
  dplyr::rename("Percent species 2" = Percent) %>%
  dplyr::select(gene,
                species1,
                `Percent species 1`,
                species2,
                `Percent species 2`,
                p.value)

p_values_total

```

## Percent resistance towards all tested compounds
This table presents the percent resistance towards the compounds sulfamethoxazole (SMX), trimethoprim (TMP), ciprofloxacin (CIP), tetracycline (TET), nalidixic acid (NAL), cefotaxime (CTX), chloramphenicol (CHL), ampicillin (AMP), and gentamicin (GEN).
```{r}
percent_am_res <- mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  # set resistant/nonresistant based on MIC cutoffs
  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  # calculate percentages
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1))

percent_am_res
```

## Resistance mechanisms per sequence type
This table presents the total resistance mechanisms identified in each sequence type (n > 2).
```{r}
mut_report %>%
  left_join(acquired_report, by = "id") %>%
  left_join(mlst_data[,c("id","ST")], by = "id") %>%
  select(-c(nt_change, id)) %>%
  gather(gene, value, -ST) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  filter(Present != 0) %>%
  mutate(Total = Present + Absent) %>%
  # filter out all STs observed less than three times
  filter(Total > 2) %>%
  rowwise() %>%
  mutate(Percent = round(Present/Total*100, 1)) %>%
  ungroup() %>%
  group_by(ST) %>%
  mutate(no_of_mechanisms = n())
```

## Resistance mechanisms per sequence type per animal species
This table presents each resistance mechanism identified in each sequence type (n > 2) per animal species.
```{r}
mut_report %>%
  left_join(acquired_report, by = "id") %>%
  left_join(mlst_data[,c("id","ST")], by = "id") %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  select(-c(nt_change, id)) %>%
  gather(gene, value, -c(ST, species)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  filter(Present != 0) %>%
  mutate(Total = Present + Absent) %>%
  # filter out all STs observed less than three times
  filter(Total > 2) %>%
  rowwise() %>%
  mutate(Percent = round(Present/Total*100, 1)) %>%
  ungroup() %>%
  group_by(ST, species) %>%
  mutate(no_of_mechanisms = n()) %>%
  arrange(ST, species)
```

## Mutations in intrinsic genes for isolates with CIP MIC < 0.12
This table presents the mutations identified in isolates with a MIC-value for ciprofloxacin below the ECOFF. Note that all mutations in all genes are listed here, not just the QRDR-located ones. 
```{r}
cip_mic_sens <- mic_data %>%
  select(id, CIP, NAL) %>%
  filter(CIP < 0.12)

acquired_genes <- acquired_report %>%
  gather(gene, value, -id) %>%
  filter(value == 1) %>%
  select(-value) %>%
  dplyr::rename("PMQR" = gene)

mut_table_fixed %>%
  filter(id %in% cip_mic_sens$id,
         gene %in% gene_names) %>%
  select(id, gene, mut) %>%
  spread(gene, mut) %>%
  left_join(acquired_genes, by = "id") %>%
  left_join(mic_data[,c("id", "CIP", "NAL")], by = "id") %>%
  left_join(isolate_data[,c("id","species","method")], by = "id")
```

## Significant differences between the amount of resistant isolates
These tables present the significant (p < 0.05) differences in isolates resistant to quinolones (table 1), and the rest of the tested compounds (table 2) for each animal species.
```{r, message=FALSE, warning=FALSE}
# create table over amount of resistant and sensitive isolates for CIP and NAL
cip_nal_res <- mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  select(compound, species, Present, Absent) %>%
  dplyr::rename("key" = compound) %>%
  split(f = .$key)

# do chi-squared test
p_values_res <- lapply(cip_nal_res, function(x) do_chisq_test(x)) %>%
  bind_rows() %>%
  # insert percentages for species 1
  left_join(percent_am_res[,c("species","compound","Percent")],
            by = c("species1" = "species",
                   "gene" = "compound")) %>%
  dplyr::rename("Percent_species1" = Percent) %>%
  # insert percentages for species 2
  left_join(percent_am_res[,c("species","compound","Percent")],
            by = c("species2" = "species",
                   "gene" = "compound")) %>%
  dplyr::rename("Percent_species2" = Percent,
                "Compound" = gene) %>%
  select(Compound,
         species1,
         Percent_species1,
         species2,
         Percent_species2,
         p.value)

p_values_res

# dataset for all other compounds
rest_res <- mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  select(compound, species, Present, Absent) %>%
  dplyr::rename("key" = compound) %>%
  split(f = .$key)

p_values_res_rest <- lapply(rest_res, function(x) do_chisq_test(x)) %>%
  bind_rows() %>%
  # insert percentages for species 1
  left_join(percent_am_res[,c("species","compound","Percent")], 
            by = c("species1" = "species",
                   "gene" = "compound")) %>%
  dplyr::rename("Percent_species1" = Percent) %>%
  # insert percentages for species 2
  left_join(percent_am_res[,c("species","compound","Percent")],
            by = c("species2" = "species",
                   "gene" = "compound")) %>%
  dplyr::rename("Percent_species2" = Percent,
                "Compound" = gene) %>%
  select(Compound,
         species1,
         Percent_species1,
         species2,
         Percent_species2,
         p.value)

p_values_res_rest
```

## Correlation between specific mutations and qnr-genes
```{r}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    base::unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    base::unlist() %>% base::unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(base::unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

col_proc <- c("gyrA","gyrB","parC","parE")

tot_mut_report <- sapply(function(i) {AddCol(df = mut_quant, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  select(-c(gyrA_0, parC_0, gyrB_0, parE_0)) %>%
  mutate_all(funs(as.numeric))

test <- mut_quant %>%
  cbind(tot_mut_report) %>%
  select(-contains("mut")) %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(-id),
            funs(as.numeric)) %>%
  mutate(qnr = ifelse(rowSums(.[,21:27]) == 0, 0, 1)) %>%
  select(-id)

cor.test(test$gyrA_S83L, test$qnr)
```

```{r, include=FALSE}
# Functions
mismatches <- function(query, ref) {
  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
                    gapOpening = 3, gapExtension = 1) %>%
    mismatchTable() %>%
    mutate(ID=names(query),
           Pos=PatternStart,
           Reference_AA=as.character(PatternSubstring),
           Sample_AA=as.character(SubjectSubstring)) %>%
    select(ID, Reference_AA, Sample_AA, Pos) %>%
    filter(Sample_AA != "X")
}

# Import sequence data
gyrA <- readAAStringSet("../data/Prokka/DNA_gyrase_GYRA_aa_sequences.fa")
parC <- readAAStringSet("../data/Prokka/Topoisomerase_IV_subunit_A_aa_sequences.fa")
parE <- readAAStringSet("../data/Prokka/DNA_topoisomerase_IV_subunit_B_aa_sequences.fa")

# Identify results from ARIBA
mut_gyrA <- mut_quant %>%
  select(id, gyrA) %>%
  mutate(entries = sapply(strsplit(.$gyrA, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(gyrA, into = as.character(c(1:max(.$entries)))) %>%
  select(-entries) %>%
  gather(key, value, `1`,`2`) %>%
  filter(is.na(value) == FALSE) %>%
  select(-key) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup()

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
gyrA_results <- bind_rows(lapply(seq_along(gyrA[-1]),
                            function(i) mismatches(gyrA[i + 1],
                                                   gyrA[1]))) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "gyrA_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, gyrA_prokka) %>%
  # control for mutation within QRDR for gyrA
  mutate(gyrA_result = gyrA_prokka %>%
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L))) %>%
  filter(gyrA_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_gyrA, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("gyrA_ariba" = value) %>%
  filter(id %not_in% ex_samples) %>%
  select(-gyrA_result)

# Identify results from ARIBA
mut_parC <- mut_quant %>%
  select(id, parC) %>%
  mutate(entries = sapply(strsplit(.$parC, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(parC, into = as.character(c(1:max(.$entries)))) %>%
  select(-entries) %>%
  gather(key, value, `1`,`2`) %>%
  filter(is.na(value) == FALSE) %>%
  select(-key) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup()

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
parC_results <- bind_rows(lapply(seq_along(parC[-1]),
                                 function(i) mismatches(parC[i + 1],
                                                        parC[1]))) %>%
  mutate(Pos = Pos + 22) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "parC_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, parC_prokka) %>%
  mutate(parC_result = parC_prokka %>%
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L))) %>%
  filter(parC_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_parC, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("parC_ariba" = value) %>%
  filter(id %not_in% ex_samples) %>%
  select(-parC_result)

# Identify results from ARIBA
mut_parE <- mut_quant %>%
  select(id, parE) %>%
  mutate(entries = sapply(strsplit(.$parE, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(parE, into = as.character(c(1:max(.$entries)))) %>%
  select(-entries) %>%
  gather(key, value, `1`,`2`) %>%
  filter(is.na(value) == FALSE) %>%
  select(-key) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup()

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
parE_results <- bind_rows(lapply(seq_along(parE[-1]),
                                 function(i) mismatches(parE[i + 1],
                                                        parE[1]))) %>%
  #mutate(Pos = Pos + 22) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "parE_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, parE_prokka) %>%
  mutate(parE_result = parE_prokka %>%
           str_extract_all("\\d+") %>% 
           map(as.integer) %>% 
           map_lgl(~ any(.x >= 366L & .x <= 523L))) %>%
  filter(parE_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_parE, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("parE_ariba" = value) %>%
  filter(id %not_in% ex_samples) %>%
  select(-parE_result)

full_report <- gyrA_results %>%
  left_join(parC_results, by = "id") %>%
  left_join(parE_results, by = "id") %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  mutate(parC_ariba = ifelse(parC_prokka == "S80I" & 
                               parC_ariba == "S58I", "S80I", parC_ariba))

full_report

```

# Figures

## Sequence types per animal group
```{r}
palette4 <- c("Livestock" = "#80b1d3",
              "Wild" = "#fb8072")

mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(animal_group = ifelse(species %in% c("Broiler","Pig"), "Livestock", "Wild")) %>%
  group_by(ST, animal_group) %>%
  dplyr::count() %>%
  group_by(ST) %>%
  mutate(total = dplyr::first(n) + dplyr::last(n)) %>%
  filter(total > 4) %>%
  ggplot(aes(factor(ST), n, fill = animal_group))+
  geom_col(color = "black")+
  labs(x = "Sequence Types")+
  scale_fill_manual(values = palette4)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.45),
        legend.title = element_blank())
```

## Resistance mechanisms per sequence type
```{r}
genes <- c(names(mut_report), names(acquired_report), "qnr")
genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")]

no_of_mlst <- mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(ST, species) %>%
  dplyr::count() %>%
  group_by(ST) %>%
  mutate(group_n = sum(n)) %>%
  filter(group_n > 3)

mlst_mechanisms_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mlst_data[c("id","ST")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, ST), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,ST)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  group_by(ST, gene, value) %>%
  dplyr::count() %>%
  filter(ST %in% no_of_mlst$ST)

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

p1 <- ggplot(mlst_mechanisms_report, aes(factor(ST), gene, size = n))+
  geom_point()+
  guides(size = FALSE)+
  labs(x = "Sequence Types",
       y = "Genes")+
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())

p2 <- ggplot(no_of_mlst, aes(factor(ST), n, fill = species))+
  geom_col(color = "black")+
  scale_fill_manual(values = palette)+
  guides(fill = FALSE)+
  theme(axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank())

plot_grid(p2, p1, nrow = 2, align = "v")

```

## Percent resistance mechanisms in total
```{r}
total_mechanisms <- mut_report %>%
  select(-nt_change) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:18]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, result, -id) %>%
  group_by(gene, result) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(result = ifelse(result == 1, "Present", "Absent")) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1),
         type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))

palette2 <- c("Acquired" = "#8dd3c7",
              "Intrinsic" = "#80b1d3")

ggplot(total_mechanisms, aes(reorder(gene, -Percent), Percent, fill = type)) +
  geom_col(color = "black") +
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.5)+
  scale_fill_manual(values = palette2) +
  labs(x = "Genes",
       y = "Percent (%) Isolates",
       fill = NULL)

```

## Percent QREC isolates with mutations in intrinsic genes
```{r}
gene_names <- names(mut_report)

gene_names <- gene_names[-c(1,2)]

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    mutate(species = factor(species,
                            levels = c("Broiler","Pig","Wild bird","Red fox")),
           key = paste0(toupper(substr(key, 1, 1)),
                        substr(key, 2, nchar(key))),
           group = if_else(key %in% c("GyrA","GyrB","ParC","ParE"), 1, 2)) %>%
  ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_manual(values = palette) +
    scale_x_discrete(labels = c("GyrA","GyrB","ParC","ParE",
                                "MarA","MarR","SoxR","RpoB",
                                "RobA")) +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

## Percent QRDR mutations per animal species
```{r}
palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  dplyr::select(-c(id, flag)) %>%
  left_join(id_data, by = "ref") %>%
  dplyr::select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples) %>%
  gather(gene, mut, -id) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  group_by(gene, mut, species) %>%
  dplyr::count() %>%
  rowwise() %>%
  mutate(Total = case_when(species == "Broiler" ~ 87,
                           species == "Pig" ~ 75,
                           species == "Red fox" ~ 52,
                           species == "Wild bird" ~ 66),
         Percent = round(n/Total * 100,1),
         lwr = round(get_binCI(n, Total)[1], 1),
         upr = round(get_binCI(n, Total)[2], 1)) %>%
  dplyr::rename("Mutation" = mut)) %>%
  ggplot(aes(reorder(Mutation, -Percent), Percent, fill = species))+
  geom_col(color = "black", position = position_dodge(0.9))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                position = position_dodge(0.9),
                width = 0.5)+
  scale_fill_manual(values = palette)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
        axis.title.x = element_blank())+
  facet_wrap(~ gene, scale = "free_x")

```

## Percent QREC isolates with Qnr genes per species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")

acquired_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(0.9)
    ) +
    scale_fill_manual(values = palette) +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.text.x = element_text(face = "italic"),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
      )
```

## Correlation plot for resistance mechanisms
```{r, fig.height=8, fig.width=7, warning=FALSE}
total_df <- acquired_report %>%
  left_join(mut_report, by = "id") %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  select(-c(nt_change, id)) %>%
  mutate_at(vars(-species),
            funs(as.numeric)) %>%
  mutate(qnr = ifelse(rowSums(.[,2:8]) > 0, 1, 0)) %>%
  {. ->> total_resistance_report} %>%
  split(., f = .$species) %>%
  lapply(., function(x) select(x, -species))

create_corr_matrix <- function(df) {
  df <- df %>%
    as.matrix %>%
    cor(method = "spearman") %>%
    as.data.frame %>%
    rownames_to_column(var = "var1") %>%
    gather(var2, value, -var1) %>%
    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
           var2 = gsub("(.*?)_R$", "\\1", var2))
  return(df)
}

corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species") %>%
  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
           gene2 = gsub("(.*?)_.+", "\\1", var2),
           var1 = gsub(".+_(.*?)", "\\1", var1),
           var2 = gsub(".+_(.*?)", "\\1", var2),
           type1 = case_when(gene1 == "gyrA" ~ 1,
                          gene1 == "gyrB" ~ 2,
                          gene1 == "parC" ~ 3,
                          gene1 == "parE" ~ 4,
                          gene1 == "marR" ~ 5,
                          gene1 == "marA" ~ 6,
                          gene1 == "rpoB" ~ 7,
                          gene1 == "soxR" ~ 8,
                          gene1 == "robA" ~ 9,
                          gene1 == "qnrA1" ~ 10,
                          gene1 == "qnrB19" ~ 11,
                          gene1 == "qnrB6" ~ 12,
                          gene1 == "qnrB60" ~ 13,
                          gene1 == "qnrS1" ~ 14,
                          gene1 == "qnrS2" ~ 15,
                          gene1 == "qnrS4" ~ 16,
                          gene1 == "qnr" ~ 17,
                          gene1 == "qepA4" ~ 18),
           type2 = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "marA" ~ 6,
                          gene2 == "rpoB" ~ 7,
                          gene2 == "soxR" ~ 8,
                          gene2 == "robA" ~ 9,
                          gene2 == "qnrA1" ~ 10,
                          gene2 == "qnrB19" ~ 11,
                          gene2 == "qnrB6" ~ 12,
                          gene2 == "qnrB60" ~ 13,
                          gene2 == "qnrS1" ~ 14,
                          gene2 == "qnrS2" ~ 15,
                          gene2 == "qnrS4" ~ 16,
                          gene2 == "qnr" ~ 17,
                          gene2 == "qepA4" ~ 18))

cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")

ggplot(corr_matrices, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient(low = "black",
                      high = "white",
                      na.value = "grey95")+
  labs(fill = NULL)+
  theme_classic()+
  theme(axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
        axis.title = element_blank())+
  coord_fixed()+
  facet_wrap(~species)
```

## Mechanisms per animal species
```{r, fig.height=10, fig.width=15, warning=FALSE, message=FALSE}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

total_df <- acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("id","marR","marA","rpoB","soxR","robA")], by = "id") %>%
  left_join(isolate_data[c("id", "species")])

names(total_df) <- c("id","qepA4","qnrA1","qnrB19",
                     "qnrB6","qnrB60","qnrS1","qnrS2",
                     "qnrS4","gyrA","gyrB","parC",
                     "parE","MarR","MarA","RpoB",
                     "SoxR","RobA","species")

col_proc <- c("gyrA","gyrB","parC","parE")

id <- total_df$id

cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  cbind(id) %>%
  gather(key, value, -id) %>%
  filter(grepl("(.*?)_0", key) == FALSE) %>%
  spread(key, value)

total_df2 <- total_df %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(cols_to_add, by = "id") %>%
  select(id, species, everything()) %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,4:10]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-id) %>%
  mutate_at(vars(-species, CIP),
            funs(as.numeric(as.character(.)))) %>%
  group_by(species, CIP) %>%
  summarise_all(funs(sum, n = n())) %>%
  dplyr::rename("n" = MarR_n) %>%
  select(-contains("_n")) %>%
  gather(gene, value, -c(species, CIP, n)) %>%
  mutate(gene2 = sub("(.*?)_(.*?)_sum", "\\1", gene),
         gene2 = sub("(.*?)_sum", "\\1", gene2),
         gene = sub("(.*?)_(.*?)_sum", "\\2", gene),
         gene = sub("(.*?)_sum", "\\1", gene),
         type = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "MarR" ~ 5,
                          gene2 == "MarA" ~ 6,
                          gene2 == "RpoB" ~ 7,
                          gene2 == "SoxR" ~ 8,
                          gene2 == "RobA" ~ 9,
                          gene2 == "qnr" ~ 10,
                          gene2 == "qepA4" ~ 11),
         perc = value/n*100)

p1 <- ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
  geom_tile()+
  scale_fill_viridis(option = "A",
                     begin = 0.1,
                     guide = guide_colorbar(barheight = 10,
                                            barwidth = 0.8,
                                            label.position = "left",
                                            label.hjust = 0.5,
                                            label.vjust = 0.3,
                                            ticks = TRUE))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        strip.text = element_text(size = 12),
        legend.position = "left",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.margin = unit(c(0,0,0,0), "cm"))+
  labs(y = "CIP MIC")+
  facet_wrap(~species, ncol = 1)+
  coord_fixed()

nplot_data <- total_df %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(species, CIP) %>%
  dplyr::count() %>%
  ungroup()

n_names <- c("Broiler" = "n",
             "Pig" = "n",
             "Red fox" = "n",
             "Wild bird" = "n")

p2 <- ggplot(nplot_data, aes(factor(CIP), n))+
  geom_col()+
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 12, face = "italic"),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        plot.margin = unit(c(0,0,-10,0), "cm"))+
  facet_wrap(~species, ncol = 1, labeller = as_labeller(n_names))+
  coord_flip()

p1g <- ggplotGrob(p1)
p2g <- ggplotGrob(p2)

g <- cbind(p1g, p2g, size = "first")

grid.arrange(g)
```

## MIC-plots
```{r}
total_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  dplyr::select(-nt_change) %>%
  mutate_at(vars(-id),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,12:18]) > 0, 1, 0)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  select(-ref) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(present,CIP) %>%
  dplyr::count() %>%
  ungroup() %>%
  rowwise() %>%
  mutate(present = ifelse(present == "", "None", present),
         total = 280,
         percent = round(n/total * 100, 1),
         lwr = round(get_binCI(n, total)[1], 1),
         upr = round(get_binCI(n, total)[2], 1)) %>%
  ungroup() %>%
  mutate(entries = sapply(strsplit(.$present, ","),
                          FUN = function(x) {length(x)}))

genes <- c(names(mut_report), names(acquired_report), "qnr")
genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")]

n_plot <- ggplot(total_report, aes(reorder(present, entries), percent, fill = factor(CIP)))+
  geom_col(color = "black")+
  labs(y = "Percent (%) isolates")+
  scale_fill_viridis(discrete = TRUE)+
  guides(fill=guide_legend(ncol = 10))+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top",
        legend.spacing.x = unit(0.2, "cm"),
        legend.justification = "center")

dot_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  select(-nt_change) %>%
  mutate_at(vars(-id),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,12:18]) > 0, 1, 0)) %>%
  select(-c(qnrA1,qnrB19,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,CIP)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  mutate(entries = sapply(strsplit(.$present, ","),
                          FUN = function(x) {length(x)}))

dot_plot <- ggplot(dot_report, aes(reorder(present, entries), gene, fill = gene))+
  geom_point(size = 2)+
  scale_y_discrete(limits = c("gyrA","parC","parE","marA","marR","rpoB","soxR","qepA4","qnr"))+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        axis.ticks.x = element_blank())

plot_grid(n_plot, dot_plot, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
```

## Additional resistance levels per animal species
```{r}
mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1),
         species = factor(species, levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
  ggplot(aes(reorder(compound, -Percent), Percent, fill = species))+
  geom_col(color = "black")+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6)+
  guides(fill = FALSE)+
  scale_fill_manual(values = palette)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))+
  facet_wrap(~species)
```

## CIP/NAL MIC-values per animal species
```{r}
mic_data[,c("id","CIP")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(CIP,species) %>%
  dplyr::count() %>%
  ggplot(aes(factor(CIP), n, fill = species))+
  geom_col(color = "black", position = position_fill())+
  labs(y = "Percent isolates",
       x = "CIP MIC")+
  scale_fill_manual(values = palette)+
  scale_y_continuous(labels = percent_format())+
  theme(legend.title = element_blank())

mic_data[,c("id","NAL")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(NAL,species) %>%
  dplyr::count() %>%
  ggplot(aes(factor(NAL), n, fill = species))+
  geom_col(color = "black", position = position_fill())+
  labs(y = "Percent isolates",
       x = "NAL MIC")+
  scale_fill_manual(values = palette)+
  scale_y_continuous(labels = percent_format())+
  theme(legend.title = element_blank())

```