### 2019-05-05 notebooks/fastqc_analysis.Rmd

---
title: "Fastqc Analysis of Raw Reads"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Dependencies
This analysis depends on the following packages:

- fastqcr

- dplyr

- ggplot2

- tidyr

- viridis

- ggsci

- scales

```{r, include = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(fastqcr,dplyr,ggplot2,tidyr,viridis,ggsci,scales)
```

```{r, include = FALSE}
input_dir <- "C:/Users/VI1511/OneDrive - Veterinærinstituttet/R/R_Projects/qrec_wgs1/data/fastqc"
```

# Functions
```{r}
# Function that lists all the file names in the input folder
file_names <- function(filepath, folder) {
  files <- list.files(path = paste0(filepath, "/", folder), pattern = "_fastqc.zip")
  return(files)
}

# Function that searches recursively for all filenames with fastqc.zip
file_names_recursive <- function(filepath) {
  files <- list.files(path = filepath, pattern = "_fastqc.zip", recursive = TRUE)
  return(files)
}

# Lists folders in the input folder
folder_names <- function(filepath) {
  folders <- list.files(path = filepath)
  return(folders)
}

# Identifies the names of the files and groups them accoring to their respective folders
get_grouping_variable <- function(path, folder) {
  files <- file_names(path, folder)
  df <- data.frame(files = files, group = folder, stringsAsFactors = FALSE)
  df$files <- gsub("(.*?)_fastqc.zip", "\\1", df$files)
  colnames(df) <- c("ref","group")
  return(df)
}

# Creates a data frame with grouping information of the reports
create_group_df <- function(path) {
  folders <- folder_names(path)
  if (length(folders) > 1) {
    df <- lapply(folders, function(folder) get_grouping_variable(path, folder))
    df <- bind_rows(df)
  } else {
    df <- get_grouping_variable(path, folders)
  }
  return(df)
}

# Function that filter out duplicated counts
filter_counts <- function(df) {
  df <- df %>%
    mutate(dupl = duplicated(Count)) %>%
    filter(dupl == FALSE)
  return(df)
}

# Function that prepares the sequence length data for plotting
prepare_seq_len_data <- function(list) {
  x <- split(list$sequence_length_distribution, list$sequence_length_distribution$group)
  x <- lapply(x, filter_counts)
  x <- bind_rows(x)
  return(x)
}

# Function that imports and wrangles fastqc data
get_fastqc_data <- function(filepath) {
  folders <- folder_names(filepath)

  get_files <- file_names_recursive(filepath)

  data_list <- lapply(get_files,
                      FUN = function(file) {
                        qc_read(paste0(filepath, "/", file),
                                modules = "all",
                                verbose = FALSE)
                      })
  names(data_list) <- gsub("(.*?)/(.*?)_fastqc.zip", "\\2", get_files)
  data_list <- purrr::transpose(data_list)

  data_list$sequence_length_distribution <- NULL
  data_list$kmer_content <- NULL

  list_names <- names(data_list)
  list_numbers <- 1:length(list_names)
  for (i in list_numbers) {
    assign(list_names[i], bind_rows(data_list[[i]], .id = "ref"))
  }
  df_list <- list(summary,
                  basic_statistics,
                  per_base_sequence_quality,
                  per_tile_sequence_quality,
                  per_sequence_quality_scores,
                  per_base_sequence_content,
                  per_sequence_gc_content,
                  per_base_n_content,
                  sequence_duplication_levels,
                  overrepresented_sequences,
                  adapter_content,
                  total_deduplicated_percentage)
  names(df_list) <- list_names
  df_list$basic_statistics <- df_list$basic_statistics %>%
    spread(Measure,Value) %>%
    left_join(group_df, by = "ref")
  return(df_list)
}
```

# Grouping
This function creates a data frame with the grouping information of the files, based on the folder names that hold them. Each file is specified a group name based on the folder they are in. This information is used to specify each group of files in the plots below.

```{r}
group_df <- create_group_df(input_dir)
```

# Import data
This function imports the data from the location specified.
```{r}
df_list <- get_fastqc_data(input_dir)
```

# Plotting

## Adapter content
This plot presents the adapter content for all sequences in each group. Each adapter type is specified by color. The X-axis represents the read positions, where the left side represents the start of the read and the right side the end.

```{r, fig.height=8, fig.width=12}
df_list$adapter_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  # get data ready for boxplot by gathering the columns into two columns
  gather(key,
  value,-c(ref,
  Position,
  group,
  seqlen)) %>%
  # plotting
  ggplot(aes(factor(
  Position,
  levels = unique(Position),
  ordered = TRUE
  ), value, color = key)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(
  x = "Position in Read",
  y = "Percent (%) Adapter Content",
  color = NULL,
  title = "Adapter content"
  ) +
  scale_colour_jama() +
  scale_y_continuous(limits = c(0, 20)) +
  theme_classic() +
  theme(
  axis.text.x = element_blank(),
  axis.title = element_text(size = 14),
  axis.ticks.x = element_blank(),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 12),
  legend.position = "bottom"
  ) +
  # creates separate windows for each group
  facet_wrap( ~ group, scales = "free_x", dir = "v")
```

## Per base sequence content
This plot presents the per base sequence content for all reads in each group. Each base is represented by color. 
```{r, fig.height=8, fig.width=12}
df_list$per_base_sequence_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  gather(key, value,-c(ref, Base, group, seqlen)) %>%
  ggplot(aes(factor(
  Base, levels = unique(Base), ordered = TRUE
  ), value, color = key)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(
  x = "Position in Read",
  y = "Percent (%)",
  color = NULL,
  title = "Per base sequence content"
  ) +
  theme_classic() +
  scale_color_jama() +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 12),
  legend.position = "bottom"
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x", dir = "v")
```

## Sequence duplication levels
This plot presents the duplication levels for all reads in each group. The X-axis contain "bins" of different duplication levels, and the number is a representation of how many of the same read is present. For example, the bin "3" represents the reads that have triplicates in the file.

```{r, fig.height=8, fig.width=12}
df_list$sequence_duplication_levels %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  gather(key, value,-c(ref, `Duplication Level`, group, seqlen)) %>%
  ggplot(aes(
  factor(
  `Duplication Level`,
  levels = unique(`Duplication Level`),
  ordered = TRUE
  ),
  value,
  fill = key
  )) +
  geom_boxplot(outlier.size = 0.5) +
  scale_fill_manual(values = c("#ef8a62",
  "#67a9cf")) +
  theme_classic() +
  labs(x = "Duplication Level",
  y = "Percent (%) of Sequences",
  fill = NULL,
  title = "Sequence duplication levels") +
  theme(
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 12),
  legend.position = "bottom",
  axis.text.x = element_text(
  size = 12,
  angle = 90,
  hjust = 1,
  vjust = 0.4
  )
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x")
```

## Sequence quality scores
This plot presents the sequence quality scores for all reads in each group. The X-axis represent the phred-score for each read, and the y-axis represent the amount of reads per phred-score. Higher score is better.

```{r, fig.height=8, fig.width=12}
df_list$per_sequence_quality_scores %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  ggplot(aes(factor(Quality), Count, fill = factor(Quality))) +
  geom_boxplot(outlier.size = 0.5) +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40)) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Quality",
  y = "Number of reads",
  title = "Per sequence quality scores") +
  guides(fill = FALSE) +
  theme_classic() +
  theme(
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14)
  ) +
  # creates separate windows for each group
  facet_wrap(~ group)
```

## Per sequence GC-content
This plot presents the % GC-content for all reads in each group.
```{r, fig.height=8, fig.width=12}
df_list$per_sequence_gc_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  ggplot(aes(factor(`GC Content`), Count)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(x = "GC content (%)",
  y = "Number of reads",
  title = "Per sequence GC content") +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(breaks = as.character(seq(
  from = 0, to = 100, by = 10
  ))) +
  theme_classic() +
  theme(
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14)
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x")
```

## Per base N-content
This plot presents the per base N content for all reads in each group. N's are added to the read instead of a base if the sequencer is unable to do the base call with sufficient confidence. As before, the X-axis represents the position in the read, where left is the start of the read and right is the end.
```{r, fig.height=8, fig.width=12}
df_list$per_base_n_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  ggplot(aes(factor(
  Base, levels = unique(Base), ordered = TRUE
  ), `N-Count`)) +
  geom_boxplot(fill = "#e6e6e6",
  outlier.size = 0.5) +
  labs(x = "Position in read",
  title = "Per base N content") +
  guides(fill = FALSE) +
  theme_classic() +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12)
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x", dir = "v")
```

## Total deduplicated percentage
This plot presents the total deduplicated percentage for all reads in each group.
```{r, fig.height=8, fig.width=12}
df_list$total_deduplicated_percentage %>%
  gather(key = "ref", value = `1`) %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length",
  perc = `1`) %>%
  mutate(perc = as.numeric(perc)) %>%
  ggplot(aes(factor(group), perc)) +
  geom_boxplot(fill = "#e6e6e6",
  outlier.size = 0.5) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "Group",
  y = "Total percentage of deduplicated reads",
  title = "Total deduplicated percentage") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
  axis.title = element_text(size = 14))
```

## Per base sequence quality
This plot presents the per base mean sequence quality for all reads in each group. The green line depicts the "good quality" limit of the phred-score 28. As before, the X-axis represents the position in each read.
```{r, fig.height=8, fig.width=12}
df_list$per_base_sequence_quality %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  mutate(Base = factor(Base,
  levels = unique(Base),
  ordered = TRUE)) %>%
  group_by(group) %>%
  mutate(xmax = length(unique(Base)) + 1) %>%
  ungroup() %>%
  ggplot(aes(Base,
  Mean)) +
  geom_boxplot(outlier.size = 0.4,
  fill = "#7f7f7f") +
  geom_hline(aes(yintercept = 28),
  color = "green") +
  labs(x = "Position in read",
  y = "Sequence quality",
  title = "Per base mean sequence quality") +
  scale_y_continuous(limits = c(0, 42)) +
  theme_classic() +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12)
  ) +
  # creates separate windows for each group
  facet_wrap( ~ group, scales = "free_x", dir = "v")
```

## Size of genome per sequencing method
```{r}
df_list$basic_statistics %>%
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename("tot_seq" = `Total Sequences`) %>%
  mutate(tot_seq = as.numeric(tot_seq)) %>%
  ggplot(aes(group.x, tot_seq))+
  geom_boxplot()+
  labs(y = "Total Sequences")+
  theme_classic()+
  theme(axis.title.x = element_blank())
```