### 2019-08-20 notebooks/Clustering analysis.Rmd
removed cgmlst results and added evolutionary models to the trees

---
title: "Clustering analysis and statistics"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtree)
library(ape)
library(tibble)
library(vegan)
library(distanceR)
library(impoRt)
library(readxl)
library(readr)
library(iteRate)
library(treeio)
library(adephylo)
library(phangorn)
library(RColorBrewer)
library(kableExtra)
```

```{r import data}
# Vector of excluded samples
ex_samples <- c(
  "2016-17-565-1-S",
  "2016-02-428-2-S",
  "2016-02-486-2-S",
  "2016-02-732-2-S",
  "2014-01-1741-1-S"
  )

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

"%not_in%" <- Negate("%in%")

# Isolate data
id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

mlst_data <- read.table("../data/mlst_results.txt",
                        sep = "\t",
                        header = TRUE,
                        stringsAsFactors = F)

```

# Patristic distances
Patristic distances were calculated from the core gene alignment SNP tree.
```{r fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
location_data <- read_excel("../data/st_location_data.xlsx") %>%
  select(id, Kommune, Fylke) %>%
  mutate(fyl_id = group_indices(., Fylke),
         kom_id = group_indices(., Kommune)) %>%
  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
  select(id, Location)

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE) %>%
  left_join(location_data, by = "id") %>%
  mutate(Year = substr(id, 1, 4)) %>%
  dplyr::rename("Species" = species)

iqtree <- read.iqtree("../data/Phylogeny/Roary/iqtree_roary.contree")
iqtree_clean <- fix_tip_labels(iqtree, id_data, "_pilon_spades.fasta", tree_type = "treedata")

pat_dist <- as.matrix(distTips(iqtree_clean@phylo, method = "patristic"))

iqtree_clean@phylo <- midpoint(iqtree_clean@phylo, labels = "support")

pat_dist_id <- as.data.frame(pat_dist) %>%
  rownames_to_column("isol1") %>%
  gather(isol2, value, -isol1) %>%
  filter(isol1 != isol2) %>%
  left_join(tree_metadata, by = c("isol1" = "id")) %>%
  rename("ST1" = ST,
         "Species1" = Species,
         "Location1" = Location,
         "Year1" = Year) %>%
  left_join(tree_metadata, by = c("isol2" = "id")) %>%
  rename("ST2" = ST,
         "Species2" = Species,
         "Location2" = Location,
         "Year2" = Year) %>%
  filter(value < 0.005) %>%
  mutate(selection = case_when(Species1 == Species2 & Location1 != Location2 ~ "ss",
                               Species1 != Species2 & is.na(Location1) == FALSE ~ "as")) %>%
  filter(selection %in% c("ss", "as"))

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE) %>%
  left_join(location_data, by = "id") %>%
  mutate(Year = substr(id, 1, 4)) %>%
  dplyr::rename("Species" = species) %>%
  mutate(ST_new = ifelse(ST %in% pat_dist_id$ST1, ST, "Other"))

pat_dist_bin <- as.data.frame(ifelse(pat_dist < 0.003, "< 0.003", "> 0.003"))

col_palette <- c("< 0.003" = "#bebada",
                 "> 0.003" = "grey95")

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

p <- ggtree(iqtree_clean) %<+% tree_metadata +
  geom_tiplab(aes(label = ST),
              size = 1,
              align = TRUE) +
  scale_color_manual(values = palette) +
  geom_tippoint(aes(color = Species))

gheatmap(p,
         pat_dist_bin,
         color = NULL,
         offset = 0.01,
         colnames = FALSE) +
  scale_fill_manual(values = col_palette) + 
  annotate("label", x = 0.26, y = 274, label = "Clade A") +
  annotate("label", x = 0.27, y = 262, label = "Clade B") +
  annotate("label", x = 0.40, y = 167, label = "Clade C") +
  annotate("label", x = 0.39, y = 138, label = "Clade D") +
  annotate("label", x = 0.33, y = 57, label = "Clade E") +
  annotate("label", x = 0.32, y = 31, label = "Clade F")

```

# SNP distance statistics on clades of interest
The table presents the summary statistics on the minimum SNP distance to the closest isolate per clade of interest.
```{r}
dist_data <- get_data("../data/Phylogeny/STs",
                      "gubbins_snp_dists.tab",
                      recursive = TRUE,
                      df = FALSE)

names(dist_data) <- c("Clade D", "Clade F", "Clade A", "Clade B", "Clade E", "Clade C")

dist_clean <- lapply(dist_data, function(x){
  x %>%
    dplyr::rename("isol1" = `snp-dists 0.6.3`) %>%
    gather(isol2, value, -isol1) %>%
    filter(isol1 != isol2) %>%
    group_by(isol2) %>%
    mutate(min_val = min(value))
})

summary_list <- lapply(dist_clean, function(x) round(summary(x$min_val),0)) %>%
  bind_rows()

summary_list$Type <- c("Min","1st.Q","Median","Mean","3rd.Q","Max")

summary_list %>%
  select(Type,
         `Clade A`,
         `Clade B`,
         `Clade C`,
         `Clade D`,
         `Clade E`,
         `Clade F`) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

# Maximum likelihood trees on clades of interest
Each tree is midpoint rooted for visual purposes. The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality.

Evolutionary models:

- Clade A: K2P+ASC+R2

- Clade C: K3P+ASC

- Clade D: K2P+ASC+R2

- Clade E: K3P+ASC

```{r message=FALSE, warning=FALSE}
# Import isolate information and tree metadata
location_data <- read_excel("../data/st_location_data.xlsx") %>%
  select(id, Kommune, Fylke) %>%
  mutate(fyl_id = group_indices(., Fylke),
         kom_id = group_indices(., Kommune)) %>%
  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
  select(id, Location)

# Import trees
ST10_raw <- read.iqtree("../data/Phylogeny/STs/ST10/iqtree/iqtree.contree")
ST162_A_raw <- read.iqtree("../data/Phylogeny/STs/ST162_A/iqtree/iqtree.contree")
ST355_raw <- read.iqtree("../data/Phylogeny/STs/ST355/iqtree/iqtree.contree")
ST744_raw <- read.iqtree("../data/Phylogeny/STs/ST744/iqtree/iqtree.contree")

raw_trees <- list(ST10_raw,
                  ST162_A_raw,
                  ST355_raw,
                  ST744_raw)

names(raw_trees) <- c("Clade D","Clade A","Clade E","Clade C")

# Clean tree tip labels
clean_trees <- lapply(
  raw_trees,
  function(x) fix_tip_labels(
    x,
    id_data,
    "_pilon_spades.fasta",
    tree_type = "treedata"
    )
  )

clean_trees <- lapply(clean_trees,
                      function(x) {
                        x@data$Bootstrap <- factor(
                          case_when(
                            x@data$UFboot >= 95 ~ "> 95",
                            x@data$UFboot < 95 &
                              x@data$UFboot >= 80 ~ "80 - 94",
                            x@data$UFboot < 80 &
                              x@data$UFboot >= 50 ~ "50 - 79",
                            x@data$UFboot < 50 ~ "< 50"
                          ),
                          ordered = TRUE,
                          levels = c("> 95",
                                     "80 - 94",
                                     "50 - 79",
                                     "< 50")
                        )
                        return(x)
                      })

# Specify color and shape palettes
palette_shape <- c(
  "Broiler" = 21,
  "Pig" = 22,
  "Red fox" = 23,
  "Wild bird" = 24
)

palette_color <- brewer.pal(11, "Set3")

year_val <- c(
  "2006",
  "2007",
  "2008",
  "2009",
  "2010",
  "2011",
  "2012",
  "2013",
  "2014",
  "2015",
  "2016")

names(palette_color) <- year_val

boot_palette <- c("> 95" = "#6a51a3",
              "80 - 94" = "#9e9ac8",
              "50 - 79" = "#cbc9e2",
              "< 50" = "#f2f0f7")

# Generate trees
annotate_tree(
  clean_trees$`Clade A`,
  tree_metadata,
  layout = "rectangular",
  tree_type = "treedata",
  midroot = TRUE,
  line_width = 0.5,
  label_variable = "Location",
  color_variable = "Year",
  shape_variable = "Species",
  bootstrap_var = "Bootstrap",
  clade_label_node = 27,
  clade_label = "18 SNP diff",
  cladelabel_offset = 0.001,
  bootstrap_lab = FALSE,
  tippoint_size = 4,
  label_offset = 0.0002,
  shape_palette = palette_shape,
  color_palette = palette_color,
  node_palette = boot_palette
) +
  xlim(0, 0.009) +
  ggtitle("Clade A, ST162")

annotate_tree(
  clean_trees$`Clade C`,
  tree_metadata,
  layout = "rectangular",
  tree_type = "treedata",
  midroot = TRUE,
  line_width = 0.5,
  label_variable = "Location",
  color_variable = "Year",
  shape_variable = "Species",
  bootstrap_var = "Bootstrap",
  clade_label_node = 12,
  clade_label = "50 SNP diff",
  cladelabel_offset = 0.00024,
  bootstrap_lab = FALSE,
  tippoint_size = 4,
  label_offset = 0.00007,
  shape_palette = palette_shape,
  color_palette = palette_color,
  node_palette = boot_palette
) +
  xlim(0, 0.0030) +
  ggtitle("Clade C, ST744")

annotate_tree(
  clean_trees$`Clade D`,
  tree_metadata,
  layout = "rectangular",
  tree_type = "treedata",
  midroot = TRUE,
  line_width = 0.5,
  label_variable = "Location",
  color_variable = "Year",
  shape_variable = "Species",
  bootstrap_var = "Bootstrap",
  clade_label_node = 19,
  clade_label = "10 SNP diff",
  cladelabel_offset = 0.00055,
  bootstrap_lab = FALSE,
  bootlab_size = 2,
  tippoint_size = 4,
  label_offset = 0.0001,
  shape_palette = palette_shape,
  color_palette = palette_color,
  node_palette = boot_palette
) +
  xlim(0, 0.005) +
  ggtitle("Clade D, ST10")

annotate_tree(
  clean_trees$`Clade E`,
  tree_metadata,
  layout = "rectangular",
  tree_type = "treedata",
  midroot = TRUE,
  line_width = 0.5,
  label_variable = "Location",
  color_variable = "Year",
  shape_variable = "Species",
  bootstrap_var = "Bootstrap",
  clade_label_node = 26,
  clade_label = "6 SNP diff",
  cladelabel_offset = 0.0013,
  bootstrap_lab = FALSE,
  tippoint_size = 4,
  label_offset = 0.0003,
  shape_palette = palette_shape,
  color_palette = palette_color,
  node_palette = boot_palette
) +
  xlim(0, 0.0165) +
  ggtitle("Clade E, ST355")

```

# P-value calculations on SNP distances within animal species
The table presents the non-exact p-values based on 1000 iterations of minimum SNP distances to closest isolate within animal species. The figure shows the distribution of the expected values with the observed values for each animal species denoted as an arrow. Red fox results is excluded from the plot as the observed distance values were too high to visualize.
```{r}
# Import distance data
total_dist <- read_delim("../data/snp_dists/final_Roary_gubbins_SNP_dists.tab",
                         delim = "\t") %>%
  rename("isol1" = `snp-dists 0.6.3`)

# Import isolate data
id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# Set row/column values to unique species ID
row_id <- as.data.frame(total_dist$isol1) %>%
  rename("ref" = `total_dist$isol1`) %>%
  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(id) %>%
  left_join(isolate_data[,c("id", "species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         isol1 = paste(species, n, sep = "_")) %>%
  ungroup() %>%
  select(isol1)

col_id <- as.data.frame(names(total_dist)) %>%
  rename("ref" = `names(total_dist)`) %>%
  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(id) %>%
  left_join(isolate_data[,c("id", "species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         isol1 = paste(species, n, sep = "_")) %>%
  ungroup() %>%
  select(isol1) %>%
  mutate(isol1 = ifelse(isol1 == "NA_1", "isol1", isol1))

total_dist$isol1 <- row_id$isol1
names(total_dist) <- col_id$isol1

# Create data frame for iteration analysis
run_df <- total_dist %>%
  gather(isol2, value, -isol1) %>%
  filter(isol1 != isol2) %>%
  mutate(species1 = sub("_.+", "", isol1),
         species2 = sub("_.+", "", isol2)) %>%
  filter(species1 == species2) %>%
  group_by(isol2) %>%
  mutate(min_val = min(value)) %>%
  ungroup() %>%
  select(isol1, isol2, min_val, species1, species2) %>%
  rename("value" = min_val,
         "group1" = species1,
         "group2" = species2)

# Run iterations
df <- iterate(run_df, runs = 1000, method = "median")

# Filter out the observed values
segment_data <- filter(df, iter == 0,
                       group1 %in% c("Broiler", "Pig", "Wild bird"))

# Calculate non-exact p-values
df2 <- df %>%
  group_by(group1) %>%
  mutate(test = case_when(group1 == "Broiler" & metric <= 12 ~ 1,
                          group1 == "Wild bird" & metric <= 53 ~ 1,
                          group1 == "Pig" & metric <= 212 ~ 1,
                          group1 == "Red fox" & metric <= 1388.5 ~ 1,
                          TRUE ~ 0)) %>%
  split(f = .$group1)

lapply(df2, function(x) round(sum(x$test)/length(x$test), 3)) %>%
  bind_rows() %>%
  mutate(Iterations = 1000) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot results, exclude red fox (too high values for the plot to be visible)
plot_df <- df %>%
  filter(group1 %in% c("Broiler", "Pig", "Wild bird"))

ggplot(plot_df, aes(metric)) +
  geom_histogram(
    aes(y = ..density..),
    binwidth = 0.2,
    color = "black",
    fill = "grey80"
  ) +
  stat_function(fun = dnorm,
                args = with(df, c(
                  mean = mean(metric), sd = sd(metric)
                )),
                color = "red") +
  geom_segment(
    segment_data,
    mapping = aes(
      x = metric,
      xend = metric,
      y = 0.5,
      yend = 0,
      color = group1
    ),
    arrow = arrow(length = unit(0.3, "cm"),
                  type = "closed")
  ) +
  geom_text(segment_data,
            mapping = aes(label = group1,
                          x = metric),
            y = 1,
            angle = 90) +
  guides(color = FALSE) +
  labs(x = "Min. number of SNPs to closest isolate",
       y = "Density") +
  ggtitle(label = "Distribution of expected values",
          subtitle = paste("Number of iterations: ", max(df$iter))) +
  theme(plot.title = element_text(hjust = 0)) +
  scale_color_manual(values = palette)

```

# NMDS Analysis of resistance mechanisms
Non-metric multidimensional scaling (NMDS) was used to see if the distribution of quinolone resistance mechanisms from some sequence types were more homogenous than in other sequence types.
```{r}
res_mechanisms <- read.table("../data/MDS_data.txt",
                       sep = "\t",
                       header = TRUE,
                       stringsAsFactors = FALSE) %>%
  column_to_rownames("id")

bin_dist <- dist(res_mechanisms,
                 method = "binary")

nmds_df <- metaMDS(bin_dist,
                   trymax = 200,
                   trace = FALSE) # silence the output

print(paste0("Stress: ", round(nmds_df$stress, 3)))

stressplot(nmds_df)

```