### 2019-11-17 notebooks/supplementary.Rmd
last adjustments

---
title: "Supplementary material"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    pdf_document:
      toc: yes
      toc_depth: '2'
---

```{r message = FALSE, warning=FALSE, echo=FALSE}
# Libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtree)
library(ape)
library(tibble)
library(vegan)
library(distanceR)
library(impoRt)
library(readxl)
library(readr)
library(iteRate)
library(treeio)
library(adephylo)
library(phangorn)
library(RColorBrewer)
library(kableExtra)
```

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r Import data, include = FALSE}
# Vector of excluded samples
ex_samples <- c(
  "2016-17-565-1-S",
  "2016-02-428-2-S",
  "2016-02-486-2-S",
  "2016-02-732-2-S",
  "2014-01-1741-1-S"
  )

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

"%not_in%" <- Negate("%in%")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

mic_data <- read.table("../data/new_mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

mic_data2 <- read.table("../data/new_mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_data <- read.table("../data/mlst_results.txt",
                        sep = "\t",
                        header = TRUE,
                        stringsAsFactors = F)

acquired_report <- read.table("../data/ariba/resfinder_quin_report.txt",
                              sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE)

mut_report <- read.table("../data/ariba/megares_quin_report.txt",
                              sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE)

```

\newpage
# Section 1: Supplementary tables
## Table S1: Group listing
The table presents the different phenotypic groups used in the isolate selection, and the number of isolates from each animal species in each group. The number of resistances within each group is listed in the column "No. of resistances".
\
```{r,echo=FALSE}

isolate_data2 <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mic_data2 %>%
  left_join(isolate_data2[,c("id","species")], by = "id") %>%
  select(antres, CIP, NAL, species) %>%
  group_by(antres, CIP, NAL) %>%
  mutate(n = n(),
         Group = group_indices()) %>%
  group_by(species, Group) %>%
  mutate(n_species = n(),
         new_id = paste0(species, " (", n_species, ")")) %>%
  group_by(Group) %>%
  summarise_all(list(func_paste)) %>%
  select(Group, antres, CIP, NAL, new_id, n) %>%
  dplyr::rename("No. of resistances" = antres,
                "Total isolates" = n,
                "Isolates per species" = new_id) %>%
  ungroup() %>%
  kable(longtable = TRUE, booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped","hover"),
                latex_options = c("hold_position", "repeat_header"))
```

\newpage
## Table S2: Occurrence of antimicrobial resistance among the selected quinolone resistant isolates.
The table presents the percent (%) occurrence of antimicrobial resistance among the selected isolates (_n_ = 280) identified through minimum inhibitory concentration values in the Norwegian monitoring programme from 2006 to 2017. The epidemiological cut off values used were defined by EUCAST. Azithromycin was excluded as no epidemiological cutoff value is currently available.
\

```{r, echo=FALSE, message=FALSE}
t1 <- read_delim("../data/phenotypic_resistance.txt",
           delim = "\t")

names(t1)[2] <- paste0(names(t1)[2],"\\textsuperscript{a}")
names(t1)[3] <- paste0(names(t1)[3],"\\textsuperscript{b}")
names(t1)[4] <- paste0(names(t1)[4],"\\textsuperscript{c}")
names(t1)[5] <- paste0(names(t1)[5],"\\textsuperscript{d}")
names(t1)[6] <- paste0(names(t1)[6],"\\textsuperscript{e}")

t1 %>%
  kable(booktabs = TRUE, escape = FALSE) %>%
  kable_styling(full_width = TRUE, position = "center") %>%
  footnote(alphabet = c("87 isolates",
                        "75 isolates",
                        "52 isolates",
                        "66 isolates",
                        "280 isolates"))
```

\newpage
## Table S3: Prokka reference genomes
The genomes listed below were used as references for the Prokka annotation, and downloaded with ncbi-genome-download (https://github.com/kblin/ncbi-genome-download) with the following commands:

--refseq-category reference

--assembly-level complete

--genus "Escherichia coli" bacteria
\

| Accession number           |  Information                 |
|----------------------------|------------------------------|
| GCF_000005845.2_ASM584v2   | _E. coli_ K12                |
| GCF_000008865.2_ASM886v2   | _E. coli_ O157:H7 Sakai      |
| GCF_000026345.1_ASM2634v1  | _E. coli_ IAI39              |
| GCF_000183345.1_ASM18334v1 | _E. coli_ O83:H1 NRG 857C    |
| GCF_000299455.1_ASM29945v1 | _E. coli_ O104:H4 2011C-3493 |

\newpage
## Table S4: Identified amino acid substitutions in the QRDR of GyrA, ParC and ParE.
The table presents the different amino acid substitutions identified in GyrA, ParC and ParE, their respective nucleotide changes, and percent occurrence in total for each amino acid substitution. The Reference nt column denote the nucleotide codon in the reference used from the MegaRes database, while the Contig nt denote the identified codon in the isolates.
\

| Gene | AA Substitution | Reference nt |	Contig nt	| n	  | Total | Percent |
|------|-----------------|--------------|-----------|-----|-------|---------|
| GyrA | S83L	           | TCG	        | TTG       |	204 |	280	  | 72.9    |
|	     | S83A	           | TCG	        | GCG       |	9	  | 280	  | 3.2     |
|	     | D87Y	           | GAC	        | TAC       |	10	| 280	  | 3.6     |
|	     | D87N	           | GAC	        | AAC       |	36	| 280	  | 12.9    | 
|	     | D87H	           | GAC	        | CAC       |	1	  | 280	  | 0.4     |
|	     | D87G	           | GAC	        | GGC       |	2	  | 280	  | 0.7     |
| ParC | S80R	           | AGC	        | AGA       |	2	  | 280	  | 0.7     |
|	     | S80I	           | AGC	        | ATC       |	37	| 280	  | 13.2    |
|	     | S80I	           | AGC	        | ATT       |	2	  | 280	  | 0.7     |
|	     | S57T	           | AGC	        | ACC       |	2	  | 280	  | 0.7     |
|	     | E84V	           | GAA	        | GTA       |	1	  | 280	  | 0.4     |
|	     | A56T	           | GCC	        | ACC       |	8	  | 280	  | 2.9     |
| ParE | S458A	         | TCG	        | GCG       |	10	| 280	  | 3.6     |
|	     | L488M	         | CTC	        | ATG       |	1	  | 280	  | 0.4     |
|	     | L416F	         | CTT	        | TTT       |	1	  | 280	  | 0.4     |
|	     | H516Y	         | CAC	        | TAC       |	1	  | 280	  | 0.4     |
|	     | D475E	         | GAC	        | GAA       |	15	| 280	  | 5.4     |
|	     | D463N	         | GAT	        | AAT       |	1	  | 280	  | 0.4     |
|	     | A512T	         | GCG	        | ACG       |	1	  | 280	  | 0.4     |

\newpage
## Table S5: QRDR substitutions in GyrA, ParC and ParE per animal species.
The table presents each identified amino acid substitution in GyrA, ParC and ParE and their occurrence in isolates from each animal species. The percentage is based on the number of isolates per species with the mutation in question (n) in relation to the total number of isolates per animal species (Total).
\

| Gene | AA substitution | Species   | n  | Total | Percent |
|------|-----------------|-----------|----|-------|---------|
| GyrA | S83L	           | Broiler   | 83	| 87    |	95.4    |
|	     | S83L	           | Pig	     | 50	| 75    |	66.7    |
|	     | S83L	           | Red fox   | 37	| 52    |	71.2    |
|	     | S83L	           | Wild bird | 34	| 66    |	51.5    |
|	     | S83A	           | Pig	     | 4  |	75    |	5.3     |
|	     | S83A	           | Red fox	 | 2  |	52    |	3.8     |
|	     | S83A	           | Wild bird | 3  |	66    |	4.5     |
|	     | D87Y	           | Broiler	 | 2  |	87    |	2.3     |
|	     | D87Y	           | Pig	     | 2  |	75    |	2.7     |
|	     | D87Y	           | Red fox	 | 1  |	52    |	1.9     |
|	     | D87Y	           | Wild bird | 5  |	66    |	7.6     |
|	     | D87N	           | Broiler	 | 9  |	87    |	10.3    |
|	     | D87N	           | Pig	     | 5  |	75    |	6.7     |
|	     | D87N	           | Red fox	 | 6  |	52    |	11.5    |
|	     | D87N	           | Wild bird | 16 |	66    |	24.2    |
|	     | D87H	           | Broiler	 | 1  |	87    |	1.1     |
|	     | D87G	           | Red fox	 | 2  |	52    |	3.8     |
| ParC | S80R	           | Pig	     | 2  |	75    |	2.7     |
|	     | S80I	           | Broiler	 | 8  |	87    |	9.2     |
|	     | S80I	           | Pig	     | 5  |	75    |	6.7     |
|	     | S80I	           | Red fox   | 10 |	52    |	19.2    |
|	     | S80I	           | Wild bird | 16	| 66    |	24.2    |
|	     | S57T	           | Pig	     | 2  |	75    |	2.7     |
|	     | E84V	           | Red fox	 | 1  |	52    |	1.9     |
|	     | A56T	           | Pig	     | 4  |	75    |	5.3     |
|	     | A56T	           | Red fox	 | 1  |	52    |	1.9     |
|	     | A56T	           | Wild bird | 3  |	66    |	4.5     |
| ParE | S458A	         | Pig	     | 3  |	75    |	4.0     |
|	     | S458A	         | Red fox	 | 1  |	52    |	1.9     |
|	     | S458A	         | Wild bird | 6  |	66    |	9.1     |
|	     | L488M	         | Broiler	 | 1  |	87    |	1.1     |
|	     | L416F	         | Pig	     | 1  |	75    |	1.3     |
|	     | H516Y	         | Wild bird | 1  |	66    |	1.5     |
|	     | D475E	         | Broiler	 | 13 |	87    |	14.9    |
|	     | D475E	         | Red fox	 | 2	| 52    |	3.8     |
|	     | D463N	         | Pig	     | 1	| 75    |	1.3     |
|	     | A512T	         | Broiler	 | 1	| 87    |	1.1     |

\newpage
## Table S6: The most abundant (n > 2) substitutions in MarA, MarR, RpoB and SoxR.
The table presents the most abundant identified amino acid substitutions in MarA, MarR, RpoB and SoxR. The percentage is based on how many of the total isolates (Total) with substitutions in the respective gene (n) had the described substitution.
\

|Gene  | AA Substitution | n   | Total | Percent |
|------|-----------------|-----|-------|---------|
| MarA | S127N	         | 26  | 34	   | 76.5    |
|	     | L78F	           | 6	 | 34	   | 17.6    |
| MarR | G103S	         | 210 | 211   | 99.5    |
|	     | Y137H	         | 210 | 211   | 99.5    |
|	     | K62R            | 6	 | 211   | 2.8     |
|	     | S3N	           | 6	 | 211   | 2.8     |
|	     | D118N	         | 4	 | 211   | 1.9     |
| RpoB | E320D	         | 37	 | 48	   | 77.1    |
|	     | V261A	         | 4	 | 48	   | 8.3     |
| SoxR | G74R	           | 63	 | 71	   | 88.7    |
|	     | T38S	           | 36	 | 71	   | 50.7    |
|	     | I40V	           | 4	 | 71	   | 5.6     |
|	     | A111T	         | 3	 | 71	   | 4.2     |
|	     | N45D	           | 3	 | 71	   | 4.2     |

\newpage
## Table S7: Mechanisms in isolates without substitutions in QRDR
This table presents the mechanisms identified in isolates with no amino acid substitutions in either of GyrA, GyrB, ParC and/or ParE.
\
```{r,echo=FALSE}
specific_ac_genes <- acquired_report %>%
  gather(PMQR, value, -id) %>%
  filter(value == 1) %>%
  select(-value)

mut_report %>%
  # all gyrB were 0
  filter(gyrA == 0,
         parC == 0,
         parE == 0) %>%
  left_join(specific_ac_genes, by = "id") %>%
  select(-c(nt_change, id)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  dplyr::select(-c(gyrA,gyrB,parC,parE)) %>%
  kable(booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

\newpage
## Table S8: Other plasmid mediated genes
The table presents the plasmid mediated genes related to other resistance phenotypes among the isolates.
\
```{r, echo = FALSE, message=FALSE}
read_delim("../data/VAMPIR/amr_ac/acquired_gene_report.tsv",
                             delim = "\t") %>%
  gather(Gene, value, -ref) %>% 
  group_by(Gene, value) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 2)) %>%
  filter(!Gene %in% c("mdfA","qepA4")) %>%
  filter(grepl("qnr", Gene) == FALSE) %>%
  mutate(Included = if_else(grepl("mph|lnu|msr|erm",
                                  Gene,
                                  ignore.case = T) == TRUE, "No", "Yes")) %>%
    mutate(gene2 = case_when(
      grepl("catA|cmlA|floR|aadA|aph|dfrA|sul|tet|lnuF|msrE|mph|ermB|floR", Gene) == TRUE ~ paste0("\\textit{", Gene, "}"),
                           Gene == "blaCMY2" ~ "\\textit{bla}\\textsubscript{CMY-2}",
                           Gene == "blaTEM1A" ~ "\\textit{bla}\\textsubscript{TEM-1A}",
                           Gene == "blaTEM1B" ~ "\\textit{bla}\\textsubscript{TEM-1B}",
                           Gene == "blaTEM1C" ~ "\\textit{bla}\\textsubscript{TEM-1C}",
                           Gene == "blaSHV2" ~ "\\textit{bla}\\textsubscript{SHV-2}",
                           Gene == "blaSHV12" ~ "\\textit{bla}\\textsubscript{SHV-12}",
                           Gene == "blaCTXM1" ~ "\\textit{bla}\\textsubscript{CTX-M-1}",
                           Gene == "blaCTXM15" ~ "\\textit{bla}\\textsubscript{CTX-M-15}",
                           Gene == "blaCTXM32" ~ "\\textit{bla}\\textsubscript{CTX-M-32}",
                           Gene == "blaCTXM55" ~ "\\textit{bla}\\textsubscript{CTX-M-55}",
                           Gene == "aac3IIa" ~ "\\textit{aac(3')-IIa}",
                           Gene == "aac3IId" ~ "\\textit{aac(3')-IId}")) %>%
  arrange(-Percent) %>%
  select(gene2, everything(), -Gene) %>%
  rename("Gene" = gene2) %>%
  kable(booktabs = TRUE, escape = FALSE) %>%
  kable_styling()

```

\newpage

## Table S9: Co-resistance
The table presents the number of isolates with each combination of plasmid mediated resistance genes, excluding combinations represented by only a single isolate.

```{r, message=FALSE}

merged_plasmid_genes <- read_delim("../merged_acquired_report.txt",
                                   delim = ",") %>%
  select(-c("mph","msrE","ermB","lnuF"))

mic_data2 <- read_delim("../data/new_mic_data.txt",
                       delim = "\t") %>%
  filter(!id %in% ex_samples) %>%
  select(-antres) %>%
  mutate(SMX_R = if_else(SMX > 64, 1, 0),
         TMP_R = if_else(TMP > 2, 1, 0),
         CIP_R = if_else(CIP > 0.06, 1, 0),
         TET_R = if_else(TET > 8, 1, 0),
         NAL_R = if_else(NAL > 16, 1, 0),
         CTX_R = if_else(CTX > 0.25, 1, 0),
         CHL_R = if_else(CHL > 16, 1, 0),
         AMP_R = if_else(AMP > 8, 1, 0),
         GEN_R = if_else(GEN > 2, 1, 0)) %>%
  select(id, contains("_R"))

AM_data <- mic_data2 %>%
  mutate(quin_R = if_else(CIP_R == 1 | NAL_R == 1, 1, 0)) %>%
  select(-c(CIP_R, NAL_R)) %>%
  mutate(antres = rowSums(.[,2:9])) %>%
  mutate(MDR = if_else(antres >= 3, 1, 0)) %>%
  left_join(isolate_data[,c("id","species")]) %>%
  left_join(mic_data, by = "id") %>%
  left_join(merged_plasmid_genes)

AM_data[,c(23:39)] %>%
  group_by_all() %>%
  mutate(group = group_indices(),
         n = length(group)) %>%
  ungroup() %>%
  mutate(PMQR = if_else(qnr == 1 | qepA == 1, 1, 0)) %>%
  gather(gene, value, -c(n, PMQR, group)) %>%
  filter(value == 1) %>%
  mutate(gene2 = case_when(gene %in% c("catA","cmlA","floR",
                                       "qepA","aadA","aph",
                                       "dfrA","qnr","sul",
                                       "tet") ~ paste0("\\textit{", gene, "}"),
                           gene == "blaCMY" ~ "\\textit{bla}\\textsubscript{CMY}",
                           gene == "blaTEM" ~ "\\textit{bla}\\textsubscript{TEM}",
                           gene == "blaSHV" ~ "\\textit{bla}\\textsubscript{SHV}",
                           gene == "blaCTXM" ~ "\\textit{bla}\\textsubscript{CTX-M}",
                           gene == "aac3II" ~ "\\textit{aac(3')-II}")) %>%
  group_by(PMQR, group) %>%
  summarise_all(list(func_paste)) %>%
  ungroup() %>%
  mutate(PMQR = if_else(PMQR == 1, "PMQR positive", "PMQR negative"),
         n = as.numeric(n)) %>%
  filter(n > 1) %>%
  select(PMQR, gene2, n) %>%
  mutate(gene_n = sapply(strsplit(gene2,","),FUN=function(x){length(x[x!="Null"])})) %>%
  arrange(PMQR, gene_n) %>%
  select(-c(PMQR, gene_n)) %>%
  rename("Combination of genes" = gene2,
         "Number of isolates" = n) %>%
  kable(booktabs = TRUE, escape = FALSE) %>%
  kable_styling() %>%
  pack_rows("PMQR Negative", 1,33, latex_gap_space = "1em") %>%
  pack_rows("PMQR Positive", 34,43, latex_gap_space = "1em")

```

\newpage
## Table S10: SNP distance statistics on isolates from the same species.
The table presents the summary statistics on the minimum SNP distance to the closest isolate from same animal species.
\

| Statistic | Broiler | Pig   | Red fox | Wild bird |
|-----------|---------|-------|---------|-----------|
| Min	      | 0	      | 0	    | 0	      | 0         |
| 1st.Q     | 2	      | 2	    | 9	      | 1         | 
| Median    | 12	    | 212	  | 1388    | 53        | 
| Mean	    | 3193	  | 4310  | 7210    | 5323      |
| 3rd.Q	    | 52	    | 3471  | 12945	  | 7922      |
| Max	      | 75534	  | 47826 | 46955	  | 33629     |

\newpage
# Section 2: Supplementary figures

```{r fig.width=12, fig.height=12, warning=FALSE, message=FALSE,echo=FALSE}
location_data <- read_excel("../data/st_location_data.xlsx") %>%
  select(id, Kommune, Fylke) %>%
  mutate(fyl_id = group_indices(., Fylke),
         kom_id = group_indices(., Kommune)) %>%
  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
  select(id, Location)

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE) %>%
  left_join(location_data, by = "id") %>%
  mutate(Year = substr(id, 1, 4)) %>%
  dplyr::rename("Species" = species)

iqtree <- read.iqtree("../data/Phylogeny/Roary/iqtree_roary.contree")
iqtree_clean <- fix_tip_labels(iqtree, id_data, "_pilon_spades.fasta", tree_type = "treedata")

pat_dist <- as.matrix(distTips(iqtree_clean@phylo, method = "patristic"))

iqtree_clean@phylo <- midpoint(iqtree_clean@phylo, labels = "support")

pat_dist_id <- as.data.frame(pat_dist) %>%
  rownames_to_column("isol1") %>%
  gather(isol2, value, -isol1) %>%
  filter(isol1 != isol2) %>%
  left_join(tree_metadata, by = c("isol1" = "id")) %>%
  rename("ST1" = ST,
         "Species1" = Species,
         "Location1" = Location,
         "Year1" = Year) %>%
  left_join(tree_metadata, by = c("isol2" = "id")) %>%
  rename("ST2" = ST,
         "Species2" = Species,
         "Location2" = Location,
         "Year2" = Year) %>%
  filter(value < 0.005) %>%
  mutate(selection = case_when(Species1 == Species2 & Location1 != Location2 ~ "ss",
                               Species1 != Species2 & is.na(Location1) == FALSE ~ "as")) %>%
  filter(selection %in% c("ss", "as"))

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE) %>%
  left_join(location_data, by = "id") %>%
  mutate(Year = substr(id, 1, 4)) %>%
  dplyr::rename("Species" = species) %>%
  mutate(ST_new = ifelse(ST %in% pat_dist_id$ST1, ST, "Other"))

pat_dist_bin <- as.data.frame(ifelse(pat_dist < 0.003, "< 0.003", "> 0.003"))

col_palette <- c("< 0.003" = "#bebada",
                 "> 0.003" = "grey95")

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

p <- ggtree(iqtree_clean) %<+% tree_metadata +
  geom_tiplab(aes(label = ST),
              size = 1,
              align = TRUE) +
  scale_color_manual(values = palette) +
  geom_tippoint(aes(color = Species))

gheatmap(p,
         pat_dist_bin,
         color = NULL,
         offset = 0.01,
         colnames = FALSE) +
  scale_fill_manual(values = col_palette) + 
  annotate("label", x = 0.26, y = 274, label = "Clade A") +
  annotate("label", x = 0.27, y = 262, label = "Clade B") +
  annotate("label", x = 0.40, y = 167, label = "Clade C") +
  annotate("label", x = 0.39, y = 138, label = "Clade D") +
  annotate("label", x = 0.33, y = 57, label = "Clade E") +
  annotate("label", x = 0.32, y = 31, label = "Clade F")

```

## Figure S1: Clade selection based on patristic distances
The figure presents the overall phylogenetic tree (left) and the calculated patristic distances (right) as less than (<) or higher than (>) the specified cutoff value (0.003). The denoted clades were further investigated with phylogenetic analyses. Patristic distances were calculated from the core gene SNP tree.

\newpage

```{r echo=FALSE}
res_mechanisms <- read.table("../data/MDS_data.txt",
                       sep = "\t",
                       header = TRUE,
                       stringsAsFactors = FALSE) %>%
  column_to_rownames("id")

bin_dist <- dist(res_mechanisms,
                 method = "binary")

nmds_df <- metaMDS(bin_dist,
                   trymax = 200,
                   trace = FALSE) # silence the output

stressplot(nmds_df)

```

## Figure S2: Stressplot
Non-metric multidimensional scaling (NMDS) was used to see if the distribution of quinolone resistance mechanisms from some sequence types were more homogenous than in other sequence types. The stressplot visualizes the goodness of fit.

\newpage

```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
knitr::include_graphics("../figures/corr_chrom_vs_plasmid.png")
```

## Figure S3: Correlations between chromosomal genes and plasmid mediated genes
The figure presents the Pearson correlation values for amino acid substitutions in GyrA, ParC, ParE, MarA, MarR, RpoB, and SoxR against all  plasmid mediated genes identified. Grey = negative correlation, green = positive correlation. Black lines denote 95% confidence intervals. Red stars denote significant correlations. The plasmid mediated genes are grouped based on gene family: _dfrA_ = _dfrA1_, _dfrA5_, _dfrA8_, _dfrA12_, _dfrA14_, and _dfrA17_. _tet_ = _tetA_, _tetB_, and _tetD_. _sul_ = _sul1 - 3_. _aph_ = _aph3Ia_, _aph3Ib_, and _aph6Id_. _aadA_ = _aadA1_, _aadA2_, _aadA5_, _aadA12_, _aadA13_, and _aadA22_. _AAC(3')-II_ = _AAC(3')-IIa_ and _AAC(3')-IId_. _bla_~TEM~ = _bla_~TEM-1-A~ – _bla_~TEM-1-C~. _bla_~SHV~ = _bla_~SHV-2~ and _bla_~SHV-12~. _bla_~CTX-M~ = _bla_~CTX-M-1~, _bla_~CTX-M-15~, _bla_~CTX-M-32~, and _bla_~CTX-M-55~. _qnr_ = _qnrA1_, _qnrB19_, _qnrS1_, _qnrS2_, and _qnrS4_.

\newpage

```{r echo=FALSE}
# Import distance data
total_dist <- read_delim("../data/snp_dists/final_Roary_gubbins_SNP_dists.tab",
                         delim = "\t") %>%
  rename("isol1" = `snp-dists 0.6.3`)
# Set row/column values to unique species ID
row_id <- as.data.frame(total_dist$isol1) %>%
  rename("ref" = `total_dist$isol1`) %>%
  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(id) %>%
  left_join(isolate_data[,c("id", "species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         isol1 = paste(species, n, sep = "_")) %>%
  ungroup() %>%
  select(isol1)
col_id <- as.data.frame(names(total_dist)) %>%
  rename("ref" = `names(total_dist)`) %>%
  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(id) %>%
  left_join(isolate_data[,c("id", "species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         isol1 = paste(species, n, sep = "_")) %>%
  ungroup() %>%
  select(isol1) %>%
  mutate(isol1 = ifelse(isol1 == "NA_1", "isol1", isol1))
total_dist$isol1 <- row_id$isol1
names(total_dist) <- col_id$isol1
# Create data frame for iteration analysis
run_df <- total_dist %>%
  gather(isol2, value, -isol1) %>%
  filter(isol1 != isol2) %>%
  mutate(species1 = sub("_.+", "", isol1),
         species2 = sub("_.+", "", isol2)) %>%
  filter(species1 == species2) %>%
  group_by(isol2) %>%
  mutate(min_val = min(value)) %>%
  ungroup() %>%
  select(isol1, isol2, min_val, species1, species2) %>%
  rename("value" = min_val,
         "group1" = species1,
         "group2" = species2)
# Run iterations
df <- iterate(run_df, runs = 1000, method = "median") 
# Filter out the observed values
segment_data <- filter(df, iter == 0,
                       group1 %in% c("Broiler", "Pig", "Wild bird"))
# Calculate non-exact p-values
df2 <- df %>%
  group_by(group1) %>%
  mutate(test = case_when(group1 == "Broiler" & metric <= 12 ~ 1,
                          group1 == "Wild bird" & metric <= 53 ~ 1,
                          group1 == "Pig" & metric <= 212 ~ 1,
                          group1 == "Red fox" & metric <= 1388.5 ~ 1,
                          TRUE ~ 0)) %>%
  split(f = .$group1)
# Plot results, exclude red fox (too high values for the plot to be visible)
plot_df <- df %>%
  filter(group1 %in% c("Broiler", "Pig", "Wild bird"))

ggplot(plot_df, aes(metric)) +
  geom_histogram(
    aes(y = ..density..),
    binwidth = 0.2,
    color = "black",
    fill = "grey80"
  ) +
  stat_function(fun = dnorm,
                args = with(df, c(
                  mean = mean(metric), sd = sd(metric)
                )),
                color = "red") +
  geom_segment(
    segment_data,
    mapping = aes(
      x = metric,
      xend = metric,
      y = 0.5,
      yend = 0,
      color = group1
    ),
    arrow = arrow(length = unit(0.3, "cm"),
                  type = "closed")
  ) +
  geom_text(segment_data,
            mapping = aes(label = group1,
                          x = metric),
            y = 1,
            angle = 90) +
  guides(color = FALSE) +
  labs(x = "SNP distance",
       y = "Density") +
  ggtitle(label = "Distribution of expected values",
          subtitle = paste("Number of iterations: ", max(df$iter))) +
  theme(plot.title = element_text(hjust = 0)) +
  scale_color_manual(values = palette)
```

## Figure S4: Iteration analysis results
The figure presents the observed median minimum SNP distance values for each animal species (coloured arrows) in relation to the expected values from the iteration analysis. Red fox results is excluded from the plot as the observed distance values were too high to visualize. The figure shows that isolates from broilers are more closely aggregated in the phylogenetic tree than what is randomly expected.

\newpage

```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeA_ST162A.png")
```

## Figure S5: Maximum likelihood tree, Clade A
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K2P+ASC+R2. Shared genome: 88.4%.

\newpage

```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeC_ST744.png")
```

## Figure S6: Maximum likelihood tree, Clade C
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K3P+ASC. Shared genome: 88.4%.

\newpage

```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeD_ST10.png")
```

## Figure S7: Maximum likelihood tree, Clade D
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K2P+ASC+R2. Shared genome: 87.2%.

\newpage

```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeE_ST355.png")
```

## Figure S8: Maximum likelihood tree, Clade E
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K3P+ASC. Shared genome: 90.2%.

\newpage
# Section 3: Metadata and reference sequences

## 3.1: Isolate information and metadata
Metadata for all 280 included isolates can be downloaded here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/total_data.xlsx

## 3.2: FastQC analysis results
Raw read quality control results can be accessed here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/notebooks/fastqc_analysis.html

## 3.3: Mash screen results
Contaminant screening results can be accessed here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/notebooks/mash_analysis.html

## 3.4: QRDR determination of GyrB, ParC and ParE
The K12 reference sequence on the whole protein is listed as the first entry in each file. The rest of the fasta entries are taken from Jiménez Gómez et al. 2004 [1].

GyrB:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/gyrB_QRDR_ref.fasta

ParC:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/parC_QRDR_ref.fasta

ParE:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/parE_QRDR_ref.fasta

## 3.5: ARIBA flag selection
Data on ARIBA flag selection can be downloaded here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/ariba_flag_selection.xlsx

## 3.6: MEGARes and ResFinder reference sequences

MEGARes:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/total_megares_references.fa

ResFinder:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/total_resfinder_references.fa

## 3.7: QUAST results
Assembly metrics can be accessed here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/notebooks/assembly_metrics.html

\newpage
# References

[1] P.A. JG, J.E. G de los R, A. RM, P.  de PR, R. GA. 2004. Molecular basis of quinolone resistance in Escherichia coli from wild birds. Canadian Journal of Veterinary Research 68:229–231